---
title: "Report"
author: "Baoyi Shi"
output: pdf_document
header-includes:
- \usepackage{caption}
---
```{r echo=FALSE}
causal_mediation <- function(data, method = c("delta", "bootstrap", "simulation"),
                             nboot = 1000, conf = 0.95, nsims = 1000,
                             outcome, treatment, mediator, covariates,
                             vecc = NULL, interaction = TRUE, event = NULL,
                             m_star = 0, a_star = 0, a = 1,
                             mreg = c("linear", "logistic"),
                             yreg = c("linear", "logistic", "loglinear", "poisson",
                                      "quasipoisson", "negbin", "coxph", "aft_exp", "aft_weibull")) {

  if (is.null(covariates) & !is.null(vecc)) {
    warning("Incompatible arguments")
  } else if (!is.null(covariates) & is.null(vecc)) {
    vecc <- colMeans(as.data.frame(data[, covariates]))
  }

  formulas <- create_formulas(outcome = outcome, treatment = treatment, mediator = mediator,
                              covariates = covariates, interaction = interaction, event = event,
                              mreg = mreg, yreg = yreg)

  regressions <- run_regressions(formulas = formulas, outcome = outcome, treatment = treatment,
                                 mediator = mediator, covariates = covariates, interaction = interaction,
                                 event = event, mreg = mreg, yreg = yreg, data = data)

  coef <- get_coef(regressions = regressions, outcome = outcome, treatment = treatment,
                   mediator = mediator, covariates = covariates, interaction = interaction,
                   event = event, mreg = mreg, yreg = yreg)

  n <- nrow(data)

  betas <- coef$betas

  thetas <- coef$thetas

  variance <- coef$variance

  vcov_thetas <- coef$vcov_thetas

  vcov_betas <- coef$vcov_betas

  vcov_block <- coef$vcov_block

  effect_estimates <- bootstrap_step(data = data, indices = c(1:nrow(data)), outcome = outcome,
                                          treatment = treatment, mediator = mediator, covariates = covariates,
                                          vecc = vecc, interaction = interaction, event = event,
                                          m_star = m_star, a_star = a_star, a = a,
                                          mreg = mreg, yreg = yreg)

###############################Delta method: 3 way and 4 way decomposition################################

  if (method == "delta") {

    theta0 <- "x1"

    theta1 <- "x2"

    theta2 <- "x3"

    theta3 <- ifelse(interaction, paste0("x", length(thetas)), "0")

    beta0 <- paste0("x", length(thetas)+1)

    beta1 <- paste0("x", length(thetas)+2)

    XC <- ifelse(length(vecc)  > 0,
                 paste0("x", length(thetas) + 2 + 1:length(vecc), "*", "vecc_", 1:length(vecc),
                        collapse = "+"),
                 "0")

    if (yreg == "linear") {

      if (mreg == "linear") {

        cde_formula <- stringr::str_replace_all(
          paste0("(", theta1, "+", theta3, "*m_star)*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnde_formula <- stringr::str_replace_all(
          paste0("(", theta1, "+", theta3, "*(", beta0, "+", beta1,
                 "*a_star+", XC, "))*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnde_formula <- stringr::str_replace_all(
          paste0("(", theta1, "+", theta3, "*(", beta0, "+", beta1,
                 "*a+", XC, "))*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnie_formula <- stringr::str_replace_all(
          paste0("(", theta2, "*", beta1, "+", theta3, "*", beta1,
                 "*a_star)*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnie_formula <- stringr::str_replace_all(
          paste0("(", theta2, "*", beta1, "+", theta3, "*", beta1,
                 "*a)*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))


        intref_formula <- stringr::str_replace_all(
          paste0(theta3, "*(", beta0, "+", beta1, "*a_star", "+", XC,
                 "-m_star)*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        intmed_formula <- stringr::str_replace_all(
          paste0(theta3, "*", beta1, "*(a-a_star)^2"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

      } else if (mreg=="logistic") {

        cde_formula <- stringr::str_replace_all(
          paste0("(", theta1, "+", theta3, "*m_star)*(a-a_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnde_formula <- stringr::str_replace_all(
          paste0(theta1, "*(a-a_star)+(", theta3, "*(a-a_star))*exp(",
                 beta0, "+", beta1, "*a_star+", XC, ")/(1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, "))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnde_formula <- stringr::str_replace_all(
          paste0(theta1, "*(a-a_star)+(", theta3, "*(a-a_star))*exp(",
                 beta0, "+", beta1, "*a+", XC, ")/(1+exp(",
                 beta0, "+", beta1, "*a+", XC, "))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnie_formula <- stringr::str_replace_all(
          paste0("(", theta2, "+", theta3, "*a_star)*(exp(", beta0,
                 "+", beta1, "*a+", XC, ")/(1+exp(", beta0, "+",
                 beta1, "*a+", XC, "))-exp(", beta0, "+", beta1,
                 "*a_star+", XC, ")/(1+exp(", beta0, "+", beta1,
                 "*a_star+", XC, ")))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnie_formula <- stringr::str_replace_all(
          paste0("(", theta2, "+", theta3, "*a)*(exp(", beta0,
                 "+", beta1, "*a+", XC, ")/(1+exp(", beta0, "+",
                 beta1, "*a+", XC, "))-exp(", beta0, "+", beta1,
                 "*a_star+", XC, ")/(1+exp(", beta0, "+", beta1,
                 "*a_star+", XC, ")))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        intref_formula <- stringr::str_replace_all(
          paste0(theta3, "*(a-a_star)*(exp(", beta0, "+", beta1, "*a_star",
                 "+", XC, ")/(1+exp(", beta0, "+", beta1, "*a_star", "+", XC, "))-m_star)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))


        intmed_formula <- stringr::str_replace_all(
          paste0(theta3, "*(a-a_star)*(exp(", beta0, "+", beta1, "*a",
                 "+", XC, ")/(1+exp(", beta0, "+", beta1, "*a", "+", XC, "))-exp(",
                 beta0, "+", beta1, "*a_star",
                 "+", XC, ")/(1+exp(", beta0, "+", beta1, "*a_star", "+", XC, ")))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

      }

      te_formula <- paste0("(", tnie_formula, ")+(", pnde_formula, ")")

      pm_formula <- paste0("(", tnie_formula, ")/((", pnde_formula, ")+(", te_formula, "))")

      cde_prop_formula <- paste0("(", cde_formula, ")/(", te_formula, ")")

      intref_prop_formula <- paste0("(", intref_formula, ")/(", te_formula, ")")

      intmed_prop_formula <- paste0("(", intmed_formula, ")/(", te_formula, ")")

      pie_prop_formula <- paste0("(", pnie_formula, ")/(", te_formula, ")")

      overall_pm_formula <- paste0("((", pnie_formula, ")+(", intmed_formula, "))/(", te_formula, ")")

      overall_int_formula <- paste0("((", intref_formula, ")+(", intmed_formula, "))/(", te_formula, ")")

      overall_pe_formula <- paste0("((", intref_formula, ")+(", intmed_formula, ")+(", pnie_formula,
                                    "))/(", te_formula, ")")

      j <- length(vecc)

      if (j > 0) {

        for (formula in c("cde_formula", "pnde_formula", "tnde_formula", "pnie_formula", "tnie_formula",
                          "intref_formula", "intmed_formula", "te_formula", "pm_formula",
                          "cde_prop_formula", "intref_prop_formula", "intmed_prop_formula",
                          "pie_prop_formula", "overall_pm_formula", "overall_int_formula",
                          "overall_pe_formula")) {

          mid <- paste0("~", get(formula))

          for (i in 1:j) {

            mid <-   stringr::str_replace_all(mid,paste("vecc", i, sep = "_"),
                                              as.character(vecc[i]))

          }

          assign(formula, as.formula(mid))

        }
      }

      cde_se <- msm::deltamethod(cde_formula, thetas, vcov_thetas)

      pnde_se <- msm::deltamethod(pnde_formula, c(thetas, betas), vcov_block)

      tnde_se <- msm::deltamethod(tnde_formula, c(thetas, betas), vcov_block)

      pnie_se <- msm::deltamethod(pnie_formula, c(thetas, betas), vcov_block)

      tnie_se <- msm::deltamethod(tnie_formula, c(thetas, betas), vcov_block)

      intref_se <- msm::deltamethod(intref_formula, c(thetas, betas), vcov_block)

      intmed_se <- msm::deltamethod(intmed_formula, c(thetas, betas), vcov_block)

      te_se <- msm::deltamethod(te_formula, c(thetas, betas), vcov_block)

      pm_se <- msm::deltamethod(pm_formula, c(thetas, betas), vcov_block)

      pie_se <- pnie_se

      cde_prop_se <- msm::deltamethod(cde_prop_formula, c(thetas, betas), vcov_block)

      intref_prop_se <- msm::deltamethod(intref_prop_formula, c(thetas, betas), vcov_block)

      intmed_prop_se <- msm::deltamethod(intmed_prop_formula, c(thetas, betas), vcov_block)

      pie_prop_se <- msm::deltamethod(pie_prop_formula, c(thetas, betas), vcov_block)

      overall_pm_se <- msm::deltamethod(overall_pm_formula, c(thetas, betas), vcov_block)

      overall_int_se <- msm::deltamethod(overall_int_formula, c(thetas, betas), vcov_block)

      overall_pe_se <- msm::deltamethod(overall_pe_formula, c(thetas, betas), vcov_block)

      # 3 way decomposition effects and se's
      decomp3way <-  c(cde = unname(effect_estimates["cde"]), cde_se = cde_se,
                       pnde = unname(effect_estimates["pnde"]), pnde_se = pnde_se,
                       tnde = unname(effect_estimates["tnde"]), tnde_se = tnde_se,
                       pnie = unname(effect_estimates["pnie"]), pnie_se = pnie_se,
                       tnie = unname(effect_estimates["tnie"]), tnie_se = tnie_se,
                       te = unname(effect_estimates["te"]), te_se = te_se,
                       pm = unname(effect_estimates["pm"]), pm_se = pm_se)

      # 4 way decomposition effects and se's
      decomp4way <-  c(cde = unname(effect_estimates["cde"]), cde_se = cde_se,
                       intref = unname(effect_estimates["intref"]), intref_se = intref_se,
                       intmed = unname(effect_estimates["intmed"]), intmed_se = intmed_se,
                       pie = unname(effect_estimates["pie"]), pie_se = pie_se,
                       te = unname(effect_estimates["te"]), te_se = te_se,
                       cde_prop = unname(effect_estimates["cde_prop"]), cde_prop_se = cde_prop_se,
                       intref_prop = unname(effect_estimates["intref_prop"]), intref_prop_se = intref_prop_se,
                       intmed_prop = unname(effect_estimates["intmed_prop"]), intmed_prop_se = intmed_prop_se,
                       pie_prop = unname(effect_estimates["pie_prop"]), pie_prop_se = pie_prop_se,
                       overall_pm = unname(effect_estimates["overall_pm"]), overall_pm_se = overall_pm_se,
                       overall_int = unname(effect_estimates["overall_int"]), overall_int_se = overall_int_se,
                       overall_pe = unname(effect_estimates["overall_pe"]), overall_pe_se = overall_pe_se)

    } else if  (yreg %in% c("logistic", "loglinear", "poisson", "quasipoisson",
                            "negbin", "coxph", "aft_exp", "aft_weibull")) {

      if (mreg=="linear") {

        cde_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta3, "*m_star)*(a-a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnde_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta3, "*(", beta0, "+", beta1,
                 "*a_star+", XC, "+", theta2, "*variance))*(a-a_star)+0.5*",
                 theta3, "^2*variance*(a^2-a_star^2))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star),
                      "\\bvariance\\b" = as.character(variance)))

        tnde_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta3, "*(", beta0, "+", beta1,
                 "*a+", XC, "+", theta2, "*variance))*(a-a_star)+0.5*",
                 theta3, "^2*variance*(a^2-a_star^2))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star),
                      "\\bvariance\\b" = as.character(variance)))

        pnie_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta2, "*", beta1, "+", theta2, "*",
                 beta1, "*a_star)*(a-a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnie_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta2, "*", beta1, "+", theta3, "*",
                 beta1, "*a)*(a-a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        cde_err_formula <- stringr::str_replace_all(
          paste0("exp(", theta1, "*(a-a_star)+", theta2, "*m_star+", theta3,
                 "*a*m_star-(", theta2, "+", theta3, "*a_star)*(", beta0, "+",
                 beta1,"*a_star+", XC, ")-0.5*(", theta2, "+", theta3,
                 "*a_star)^2*variance)-exp(", theta2, "*m_star+", theta3,
                 "*a_star*m_star-(", theta2, "+", theta3, "*a_star)*(", beta0,
                 "+", beta1, "*a_star", "+", XC, ")-0.5*(",theta2, "+", theta3,
                 "*a_star)^2*variance)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star),
                      "\\bvariance\\b" = as.character(variance)))

        intref_err_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta3, "*(", beta0, "+",
                 beta1,"*a_star+", XC, "+", theta2, "*variance))*(a-a_star)+0.5*",
                 theta3, "^2*variance*(a^2-a_star^2))-1-exp(", theta1, "*(a-a_star)+",
                 theta2, "*m_star+", theta3, "*a*m_star-(", theta2, "+", theta3,
                 "*a_star)*(", beta0, "+", beta1, "*a_star+", XC, ")-0.5*(", theta2,
                 "+", theta3, "*a_star)^2*variance)+exp(", theta2, "*m_star+",
                 theta3, "*a_star*m_star-(", theta2, "+", theta3, "*a_star)*(", beta0,
                 "+", beta1,"*a_star+", XC, ")-0.5*(", theta2, "+", theta3,
                 "*a_star)^2*variance)"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star),
                      "\\bvariance\\b" = as.character(variance)))

        intmed_err_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta2, "*", beta1, "+", theta3,
                 "*(", beta0, "+", beta1,"*a_star+", "+", beta1,"*a+", XC,
                 "+", theta2, "*variance))*(a-a_star)+0.5*", theta3,
                 "^2*variance*(a^2-a_star^2))-exp((", theta2, "*", beta1,
                 "+", theta3, "*", beta1, "*a_star)*(a-a_star))-exp((",
                 theta1, "+", theta3, "*(", beta0, "+", beta1, "*a_star+", XC,
                 "+", theta2, "*variance))*(a-a_star)+0.5*", theta3,
                 "^2*variance*(a^2-a_star^2))+1"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star),
                      "\\bvariance\\b" = as.character(variance)))

      } else if (mreg=="logistic") {

        cde_rr_formula <- stringr::str_replace_all(
          paste0("exp((", theta1, "+", theta3, "*m_star)*(a-a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnde_rr_formula <- stringr::str_replace_all(
          paste0("exp(", theta1, "*(a-a_star))*(1+exp(", theta2, "+",
                 theta3, "*a+", beta0, "+", beta1, "*a_star+", XC,
                 "))/(1+exp(", theta2, "+", theta3, "*a_star+", beta0,
                 "+", beta1, "*a_star+", XC, "))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnde_rr_formula <- stringr::str_replace_all(
          paste0("exp(", theta1, "*(a-a_star))*(1+exp(", theta2, "+",
                 theta3, "*a+", beta0, "+", beta1, "*a+", XC,
                 "))/(1+exp(", theta2, "+", theta3, "*a_star+", beta0,
                 "+", beta1, "*a+", XC, "))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        pnie_rr_formula <- stringr::str_replace_all(
          paste0("(1+exp(", beta0, "+", beta1, "*a_star+", XC,
                 "))*(1+exp(", theta2, "+", theta3, "*a_star+",
                 beta0, "+", beta1, "*a+", XC, "))/((1+exp(",
                 beta0, "+", beta1, "*a+", XC, "))*(1+exp(",
                 theta2, "+", theta3, "*a_star+", beta0, "+", beta1,
                 "*a_star+", XC, ")))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        tnie_rr_formula <- stringr::str_replace_all(
          paste0("(1+exp(", beta0, "+", beta1, "*a_star+", XC,
                 "))*(1+exp(", theta2, "+", theta3, "*a+",
                 beta0, "+", beta1, "*a+", XC, "))/((1+exp(",
                 beta0, "+", beta1, "*a+", XC, "))*(1+exp(",
                 theta2, "+", theta3, "*a+", beta0, "+", beta1,
                 "*a_star+", XC, ")))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        cde_err_formula <- stringr::str_replace_all(
          paste0("exp(", theta1, "*(a-a_star)+", theta2, "*m_star+", theta3,
                 "*a*m_star)*(1+exp(", beta0, "+", beta1, "*a_star+", XC,
                 "))/(1+exp(",  beta0, "+", beta1,  "*a_star+", XC, "+",
                 theta2, "+", theta3, "*a_star))-exp(", theta2, "*m_star+",
                 theta3, "*a_star*m_star)*(1+exp(", beta0, "+", beta1,
                 "*a_star+", XC, "))/(1+exp(", beta0, "+", beta1, "*a_star+",
                 XC, "+", theta2, "+", theta3, "*a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))


        intref_err_formula <- stringr::str_replace_all(
          paste0("exp(", theta1, "*(a-a_star))*(1+exp(", beta0, "+",
                 beta1, "*a_star+", XC, "+", theta2, "+", theta3,
                 "*a))/(1+exp(", beta0, "+", beta1, "*a_star+", XC, "+",
                 theta2, "+", theta3, "*a_star))-1-exp(", theta1,
                 "*(a-a_star)+", theta2, "*m_star+", theta3, "*a*m_star)*(1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, "))*exp((", theta1, "+",
                 theta3, "*m_star)*(a-a_star))/(1+exp(", beta0, "+", beta1,
                 "*a_star+", XC, "+", theta2, "+", theta3, "*a_star))+exp(",
                 theta2, "*m_star+", theta3, "*a_star*m_star)*(1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, "))/(1+exp(", beta0, "+",
                 beta1, "*a_star+", XC, "+", theta2, "+", theta3, "*a_star))"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))

        intmed_err_formula <- stringr::str_replace_all(
          paste0("(exp(", theta1, "*(a-a_star))*(1+exp(", beta0, "+",
                 beta1, "*a+", XC, "+", theta2, "+", theta3, "*a))*(1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, ")))/((1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, "+", theta2, "+", theta3,
                 "*a_star))*(1+exp(", beta0, "+", beta1, "*a+", XC,
                 ")))-((1+exp(", beta0, "+", beta1, "*a+", XC, "+", theta2,
                 "+", theta3, "*a_star))*(1+exp(", beta0, "+", beta1,
                 "*a_star+", XC, ")))/((1+exp(", beta0, "+", beta1, "*a_star+",
                 XC, "+", theta2, "+", theta3, "*a_star))*(1+exp(", beta0, "+",
                 beta1, "*a+", XC, ")))-(exp(", theta1, "*(a-a_star))*(1+exp(",
                 beta0, "+", beta1, "*a_star+", XC, "+", theta2, "+", theta3,
                 "*a)))/(1+exp(", beta0, "+", beta1, "*a_star+", XC, "+",
                 theta2, "+", theta3, "*a_star))+1"),
          pattern = c("\\ba_star\\b" = as.character(a_star),
                      "\\ba\\b" = as.character(a),
                      "\\bm_star\\b" = as.character(m_star)))
      }

      te_rr_formula <- paste0("(", tnie_rr_formula, ")*(", pnde_rr_formula, ")")

      pm_formula <- paste0("((", pnde_rr_formula, ")*(", tnie_rr_formula, "-1))/(", te_rr_formula, "-1)")

      cde_err_prop_formula <- paste0("(", cde_err_formula, ")/(", te_rr_formula, "-1)")

      intmed_err_prop_formula <- paste0("(", intmed_err_formula, ")/(", te_rr_formula, "-1)")

      intref_err_prop_formula <-paste0("(", intref_err_formula, ")/(", te_rr_formula, "-1)")

      pie_err_prop_formula <- paste0("(", pnie_rr_formula, "-1)/(", te_rr_formula, "-1)")

      overall_pm_formula <- paste0("((", pnie_rr_formula, "-1)+(", intmed_err_formula, "))/(",
                                   te_rr_formula, "-1)")

      overall_int_formula <- paste0("((", intref_err_formula, ")+(", intmed_err_formula, "))/(",
                                    te_rr_formula, "-1)")

      overall_pe_formula <- paste0("((", intref_err_formula, ")+(", intmed_err_formula, ")+(",
                                   pnie_rr_formula,
                                   "-1))/(", te_rr_formula, "-1)")

      j <- length(vecc)

      if (j > 0) {

        for (formula in c("cde_rr_formula", "pnde_rr_formula", "tnde_rr_formula",
                          "pnie_rr_formula", "tnie_rr_formula", "cde_err_formula",
                          "intref_err_formula", "intmed_err_formula","te_rr_formula", "pm_formula",
                          "cde_err_prop_formula", "intref_err_prop_formula", "intmed_err_prop_formula",
                          "pie_err_prop_formula", "overall_pm_formula", "overall_int_formula",
                          "overall_pe_formula")) {

          mid <- paste0("~", get(formula))

          for (i in 1:j) {

            mid <-   stringr::str_replace_all(mid,paste("vecc", i, sep = "_"),
                                              as.character(vecc[i]))

          }

          assign(formula, as.formula(mid))

        }
      }

      cde_rr_se <- msm::deltamethod(cde_rr_formula, thetas, vcov_thetas)

      pnde_rr_se <- msm::deltamethod(pnde_rr_formula, c(thetas, betas), vcov_block)

      tnde_rr_se <- msm::deltamethod(tnde_rr_formula, c(thetas, betas), vcov_block)

      pnie_rr_se <- pie_err_se <- msm::deltamethod(pnie_rr_formula, c(thetas, betas), vcov_block)

      tnie_rr_se <- msm::deltamethod(tnie_rr_formula, c(thetas, betas), vcov_block)

      te_rr_se <- total_err_se <- msm::deltamethod(te_rr_formula, c(thetas, betas), vcov_block)

      pm_se <- msm::deltamethod(pm_formula, c(thetas, betas), vcov_block)

      cde_err_se <-  msm::deltamethod(cde_err_formula, c(thetas, betas), vcov_block)

      intref_err_se <-  msm::deltamethod(intref_err_formula, c(thetas, betas), vcov_block)

      intmed_err_se <-  msm::deltamethod(intmed_err_formula, c(thetas, betas), vcov_block)

      cde_err_prop_se <- msm::deltamethod(cde_err_prop_formula, c(thetas, betas), vcov_block)

      intmed_err_prop_se <- msm::deltamethod(intmed_err_prop_formula, c(thetas, betas), vcov_block)

      intref_err_prop_se <- msm::deltamethod(intref_err_prop_formula, c(thetas, betas), vcov_block)

      pie_err_prop_se <- msm::deltamethod(pie_err_prop_formula, c(thetas, betas), vcov_block)

      overall_pm_se <- msm::deltamethod(overall_pm_formula, c(thetas, betas), vcov_block)

      overall_int_se <- msm::deltamethod(overall_int_formula, c(thetas, betas), vcov_block)

      overall_pe_se <- msm::deltamethod(overall_pe_formula, c(thetas, betas), vcov_block)

      # 3 way decomposition effects and se's
      decomp3way <-  c(cde_rr = unname(effect_estimates["cde_rr"]), cde_rr_se = cde_rr_se,
                       pnde_rr = unname(effect_estimates["pnde_rr"]), pnde_rr_se = pnde_rr_se,
                       tnde_rr = unname(effect_estimates["tnde_rr"]), tnde_rr_se = tnde_rr_se,
                       pnie_rr = unname(effect_estimates["pnie_rr"]), pnie_rr_se = pnie_rr_se,
                       tnie_rr = unname(effect_estimates["tnie_rr"]), tnie_rr_se = tnie_rr_se,
                       te_rr = unname(effect_estimates["te_rr"]), te_rr_se = te_rr_se,
                       pm = unname(effect_estimates["pm"]), pm_se = pm_se)

      # 4 way decomposition effects and se's
      decomp4way <-  c(cde_err = unname(effect_estimates["cde_err"]), cde_err_se = cde_err_se,
                       intref_err = unname(effect_estimates["intref_err"]), intref_err_se = intref_err_se,
                       intmed_err = unname(effect_estimates["intmed_err"]), intmed_err_se = intmed_err_se,
                       pie_err = unname(effect_estimates["pie_err"]), pie_err_se = pie_err_se,
                       total_err = unname(effect_estimates["total_err"]), total_err_se = total_err_se,
                       cde_err_prop = unname(effect_estimates["cde_err_prop"]), cde_err_prop_se = cde_err_prop_se,
                       intref_err_prop = unname(effect_estimates["intref_err_prop"]),
                       intref_err_prop_se = intref_err_prop_se,
                       intmed_err_prop = unname(effect_estimates["intmed_err_prop"]),
                       intmed_err_prop_se = intmed_err_prop_se,
                       pie_err_prop = unname(effect_estimates["pie_err_prop"]), pie_err_prop_se = pie_err_prop_se,
                       overall_pm = unname(effect_estimates["overall_pm"]), overall_pm_se = overall_pm_se,
                       overall_int = unname(effect_estimates["overall_int"]), overall_int_se = overall_int_se,
                       overall_pe = unname(effect_estimates["overall_pe"]), overall_pe_se = overall_pe_se)

    }

    out <- list(decomp3way = decomp3way, decomp4way = decomp4way)

  }


#####################################Bootstrapping: 3 way and 4 way decomposition####################################

  if (method == "bootstrap") {

     out <- boot::boot(data = data, statistic = bootstrap_step, R = nboot,
                        outcome = outcome, treatment = treatment, mediator = mediator,
                        covariates = covariates, vecc = vecc, interaction = interaction,
                        event = event, mreg = mreg, yreg = yreg,
                        m_star = m_star, a_star = a_star, a = a)

     se <- apply(out$t, 2, sd)

     out <- effect_estimates

     for (i in 1:length(effect_estimates))
       out[paste0(names(effect_estimates)[i], "_se")] <- se[i]

     if (yreg == "linear") {

       # 3 way decomposition effects and se's
       decomp3way <-  c(cde = unname(out["cde"]), cde_se =  unname(out["cde_se"]),
                      pnde = unname(out["pnde"]), pnde_se =  unname(out["pnde_se"]),
                      tnde = unname(out["tnde"]), tnde_se =  unname(out["tnde_se"]),
                      pnie = unname(out["pnie"]), pnie_se =  unname(out["pnie_se"]),
                      tnie = unname(out["tnie"]), tnie_se =  unname(out["tnie_se"]),
                      te = unname(out["te"]), te_se =  unname(out["te_se"]),
                      pm = unname(out["pm"]), pm_se =  unname(out["pm_se"]))

       # 4 way decomposition effects and se's
       decomp4way <-  c(cde = unname(out["cde"]), cde_se = unname(out["cde_se"]),
                      intref = unname(out["intref"]), intref_se = unname(out["intref_se"]),
                      intmed = unname(out["intmed"]), intmed_se = unname(out["intmed_se"]),
                      pie = unname(out["pie"]), pie_se = unname(out["pie_se"]),
                      te = unname(out["te"]), te_se = unname(out["te_se"]),
                      cde_prop = unname(out["cde_prop"]), cde_prop_se = unname(out["cde_prop_se"]),
                      intref_prop = unname(out["intref_prop"]), intref_prop_se = unname(out["intref_prop_se"]),
                      intmed_prop = unname(out["intmed_prop"]), intmed_prop_se = unname(out["intmed_prop_se"]),
                      pie_prop = unname(out["pie_prop"]), pie_prop_se = unname(out["pie_prop_se"]),
                      overall_pm = unname(out["overall_pm"]), overall_pm_se = unname(out["overall_pm_se"]),
                      overall_int = unname(out["overall_int"]), overall_int_se = unname(out["overall_int_se"]),
                      overall_pe = unname(out["overall_pe"]), overall_pe_se = unname(out["overall_pe_se"]))

     } else if (yreg %in% c("logistic", "loglinear", "poisson", "quasipoisson",
                     "negbin", "coxph", "aft_exp", "aft_weibull")) {

       # 3 way decomposition effects and se's
       decomp3way <-  c(cde_rr = unname(out["cde_rr"]), cde_rr_se = unname(out["cde_rr_se"]),
                        pnde_rr = unname(out["pnde_rr"]), pnde_rr_se = unname(out["pnde_rr_se"]),
                        tnde_rr = unname(out["tnde_rr"]), tnde_rr_se = unname(out["tnde_rr_se"]),
                        pnie_rr = unname(out["pnie_rr"]), pnie_rr_se = unname(out["pnie_rr_se"]),
                        tnie_rr = unname(out["tnie_rr"]), tnie_rr_se = unname(out["tnie_rr_se"]),
                        te_rr = unname(out["te_rr"]), te_rr_se = unname(out["te_rr_se"]),
                        pm = unname(out["pm"]), pm_se = unname(out["pm_se"]))

       # 4 way decomposition effects and se's
       decomp4way <-  c(cde_err = unname(out["cde_err"]), cde_err_se = unname(out["cde_err_se"]),
                        intref_err = unname(out["intref_err"]), intref_err_se = unname(out["intref_err_se"]),
                        intmed_err = unname(out["intmed_err"]), intmed_err_se = unname(out["intmed_err_se"]),
                        pie_err = unname(out["pie_err"]), pie_err_se = unname(out["pie_err_se"]),
                        total_err = unname(out["total_err"]), total_err_se = unname(out["total_err_se"]),
                        cde_err_prop = unname(out["cde_err_prop"]), cde_err_prop_se = unname(out["cde_err_prop_se"]),
                        intref_err_prop = unname(out["intref_err_prop"]),
                        intref_err_prop_se = unname(out["intref_err_prop_se"]),
                        intmed_err_prop = unname(out["intmed_err_prop"]),
                        intmed_err_prop_se = unname(out["intmed_err_prop_se"]),
                        pie_err_prop = unname(out["pie_err_prop"]), pie_err_prop_se = unname(out["pie_err_prop_se"]),
                        overall_pm = unname(out["overall_pm"]), overall_pm_se = unname(out["overall_pm_se"]),
                        overall_int = unname(out["overall_int"]), overall_int_se = unname(out["overall_int_se"]),
                        overall_pe = unname(out["overall_pe"]), overall_pe_se = unname(out["overall_pe_se"]))

     }

     out <- list(decomp3way = decomp3way, decomp4way = decomp4way)

  }


###############################Simulation-based approach: 3 way and 4 way decomposition##################################

  if (method == "simulation") {

    EY0m_sim <- EY1m_sim <- EY00_sim <- EY01_sim <- EY10_sim <- EY11_sim <-c()

    for (j in 1:nsims) {

      data_sim <- data[sample(1:nrow(data), replace = TRUE),]

      regressions_sim <- run_regressions(formulas = formulas, outcome = outcome, treatment = treatment,
                                         mediator = mediator, covariates = covariates, interaction = interaction,
                                         event = event, mreg = mreg, yreg = yreg, data = data_sim)

      # counterfactual design matrices for the mediator model
      mdesign_a <- cbind(data.frame(treatment = c(rep(a,n))),data[,covariates])

      mdesign_a_star <- cbind(data.frame(treatment = c(rep(a_star,n))),data[,covariates])

      colnames(mdesign_a) <- colnames(mdesign_a_star) <- c(treatment, covariates)

      # simulate counterfactual mediators
      m_a <- predict(regressions_sim$mediator_regression, newdata = mdesign_a, type = "response")

      m_a_star <- predict(regressions_sim$mediator_regression, newdata = mdesign_a_star, type = "response")

      # counterfactual design matrices for the outcome model
      ydesign0m <- cbind(data.frame(treatment = c(rep(a_star,n)), mediator = rep(m_star, n)), data[,covariates])

      ydesign1m <- cbind(data.frame(treatment = c(rep(a,n)), mediator = rep(m_star, n)), data[,covariates])

      ydesign00 <- cbind(data.frame(treatment = c(rep(a_star,n)), mediator = m_a_star), data[,covariates])

      ydesign01 <- cbind(data.frame(treatment = c(rep(a_star,n)), mediator = m_a), data[,covariates])

      ydesign10 <- cbind(data.frame(treatment = c(rep(a,n)), mediator = m_a_star), data[,covariates])

      ydesign11 <- cbind(data.frame(treatment = c(rep(a,n)), mediator = m_a), data[,covariates])

      colnames(ydesign0m) <- colnames(ydesign1m) <- colnames(ydesign00) <- colnames(ydesign01) <-
        colnames(ydesign10) <- colnames(ydesign11) <- c(treatment, mediator, covariates)

      # simulate counterfactual mediators
      y0m <- predict(regressions_sim$outcome_regression, newdata =  ydesign0m, type = "response")

      y1m <- predict(regressions_sim$outcome_regression, newdata =  ydesign1m, type = "response")

      y00 <- predict(regressions_sim$outcome_regression, newdata =  ydesign00, type = "response")

      y01 <- predict(regressions_sim$outcome_regression, newdata =  ydesign01, type = "response")

      y10 <- predict(regressions_sim$outcome_regression, newdata =  ydesign10, type = "response")

      y11 <- predict(regressions_sim$outcome_regression, newdata =  ydesign11, type = "response")

      # calculate causal effect components
      EY0m_sim <- c(EY0m_sim, sum(y0m)/n) #E(Ya0m*)

      EY1m_sim <- c(EY1m_sim, sum(y1m)/n) #E(Ya1m*)

      EY00_sim <- c(EY00_sim, sum(y00)/n) #E(Ya0Ma0)

      EY10_sim <- c(EY10_sim, sum(y10)/n) #E(Ya1Ma0)

      EY01_sim <- c(EY01_sim, sum(y01)/n) #E(Ya0Ma1)

      EY11_sim <- c(EY11_sim, sum(y11)/n) #E(Ya1Ma1)

      }

    # linear outcome
    if (yreg == "linear") {

         cde_sim <- EY1m_sim - EY0m_sim # cde in the 3 way or 4 way decomposition

         pnde_sim <- EY10_sim - EY00_sim # pnde in the 3 way decomposition

         tnie_sim <- EY11_sim - EY10_sim # tnie in the 3 way decomposition

         tnde_sim <- EY11_sim - EY01_sim # tnde in the 3 way decomposition

         pnie_sim <- EY01_sim - EY00_sim # pnie in the 3 way decomposition

         te_sim <- tnie_sim + pnde_sim # te in the 3 way or 4 way decomposition

         pm_sim <- tnie_sim / (pnde_sim + te_sim) # pm in the 3 way decomposition

         intref_sim <- pnde_sim - cde_sim # intref in the 4 way decomposition

         intmed_sim <- tnie_sim - pnie_sim # intmed in the 4 way decomposition

         pie_sim <- pnie_sim # pie in the 4 way decomposition

         cde_prop_sim <- cde_sim/te_sim # proportion cde in the 4 way decomposition

         intref_prop_sim <- intref_sim/te_sim # proportion intref in the 4 way decomposition

         intmed_prop_sim <- intmed_sim/te_sim # proportion intmed in the 4 way decomposition

         pie_prop_sim <- pie_sim/te_sim # proportion pie in the 4 way decomposition

         overall_pm_sim <- (pie_sim + intmed_sim)/te_sim # overall pm in the 4 way decomposition

         overall_int_sim <- (intref_sim + intmed_sim)/te_sim # overall int in the 4 way decomposition

         overall_pe_sim <- (intref_sim + intmed_sim + pie_sim)/te_sim # overall pe in the 4 way decomposition

         # calculate the mean of each simulated effects
         for (effect in c("cde", "pnde", "tnde", "pnie", "tnie", "te", "pm", "intref", "intmed",
                          "pie", "cde_prop", "intref_prop", "intmed_prop", "pie_prop",
                          "overall_pm", "overall_int", "overall_pe")) {

           mid <- mean(get(paste0(effect, "_sim")))

           assign(effect, mid)

         }

         # calculate the se of each simulated effects
         for (effect_se in c("cde_se", "pnde_se", "tnde_se", "pnie_se", "tnie_se", "te_se", "pm_se",
                             "intref_se", "intmed_se", "pie_se", "cde_prop_se", "intref_prop_se",
                             "intmed_prop_se", "pie_prop_se", "overall_pm_se",
                             "overall_int_se", "overall_pe_se")) {

           mid <- sd(get(stringr::str_replace(effect_se, "_se", "_sim")))

           assign(effect_se, mid)

         }

         # 3 way decomposition effects and se's
         decomp3way <-  c(cde = cde, cde_se = cde_se,
                          pnde = pnde, pnde_se = pnde_se,
                          tnde = tnde, tnde_se = tnde_se,
                          pnie = pnie, pnie_se = pnie_se,
                          tnie = tnie, tnie_se = tnie_se,
                          te = te, te_se = te_se,
                          pm = pm, pm_se = pm_se)

         # 4 way decomposition effects and se's
         decomp4way <-  c(cde = cde, cde_se = cde_se, intref = intref, intref_se = intref_se,
                          intmed = intmed, intmed_se = intmed_se, pie = pie, pie_se = pie_se,
                          te = te, te_se = te_se, cde_prop = cde_prop, cde_prop_se = cde_prop_se,
                          intref_prop = intref_prop, intref_prop_se = intref_prop_se,
                          intmed_prop = intmed_prop, intmed_prop_se = intmed_prop_se,
                          pie_prop = pie_prop, pie_prop_se = pie_prop_se,
                          overall_pm = overall_pm, overall_pm_se = overall_pm_se,
                          overall_int = overall_int, overall_int_se = overall_int_se,
                          overall_pe = overall_pe, overall_pe_se = overall_pe_se)

         out <- list(decomp3way = decomp3way, decomp4way = decomp4way,
                     sims = data.frame(cde = cde_sim, pnde = pnde_sim, tnde = tnde_sim,
                                       pnie = pnie_sim, tnie = tnie_sim, intref = intref_sim,
                                       intmed = intmed_sim, pie_sim = pnie_sim))

       }

     # nonlinear outcome
     if (yreg != "linear") {

         cde_rr_sim <- EY1m_sim/EY0m_sim # cde in the 3 way decomposition

         pnde_rr_sim <- EY10_sim/EY00_sim # pnde in the 3 way decomposition

         tnie_rr_sim <- EY11_sim/EY10_sim # tnie in the 3 way decomposition

         tnde_rr_sim <- EY11_sim/EY01_sim # tnde in the 3 way decomposition

         pnie_rr_sim <- EY01_sim/EY00_sim # pnie in the 3 way decomposition

         te_rr_sim <- tnie_rr_sim * pnde_rr_sim # te in the 3 way or 4 way decomposition

         pm_sim <- (pnde_rr_sim * (tnie_rr_sim - 1)) / (pnde_rr_sim * tnie_rr_sim - 1) # pm in the 3 way decomposition

         cde_err_sim <- (EY1m_sim-EY0m_sim)/EY00_sim # excess RR due to cde in the 4 way decomposition

         intref_err_sim <- pnde_rr_sim - 1 - cde_err_sim # excess RR due to intref in the 4 way decomposition

         intmed_err_sim <- tnie_rr_sim * pnde_rr_sim - pnde_rr_sim - pnie_rr_sim + 1 # excess RR due to intmed in the 4 way decomposition

         pie_err_sim <- pnie_rr_sim - 1 # excess RR due to pie in the 4 way decomposition

         total_err_sim <- EY11_sim / EY00_sim - 1 # total excess RR in the 4 way decomposition

         cde_prop_sim <- cde_err_sim/total_err_sim # proportion cde

         intmed_prop_sim <- intmed_err_sim/total_err_sim # proportion intmed

         intref_prop_sim <- intref_err_sim/total_err_sim # proportion intref

         pie_prop_sim <- pie_err_sim/total_err_sim # proportion pie

         overall_pm_sim <- (pie_err_sim+intmed_err_sim)/total_err_sim # overall pm in the 4 way decomposition

         overall_int_sim <- (intref_err_sim+intmed_err_sim)/total_err_sim # overall int in the 4 way decomposition

         overall_pe_sim <- (intref_err_sim+intmed_err_sim+pie_err_sim)/total_err_sim # overall pe in the 4 way decomposition

         # calculate the mean of each simulated effects
         for (effect in c("cde_rr", "pnde_rr", "tnde_rr", "pnie_rr", "tnie_rr", "te_rr", "pm",
                          "cde_err", "intref_err", "intmed_err", "pie_err", "total_err",
                          "cde_prop", "intref_prop", "intmed_prop", "pie_prop",
                          "overall_pm", "overall_int", "overall_pe")) {

           mid <- mean(get(paste0(effect, "_sim")))

           assign(effect, mid)

         }

         # calculate the se of each simulated effects
         for (effect_se in c("cde_rr_se", "pnde_rr_se", "tnde_rr_se", "pnie_rr_se", "tnie_rr_se",
                             "te_rr_se", "pm_se", "cde_err_se", "intref_err_se", "intmed_err_se",
                             "pie_err_se", "total_err_se", "cde_prop_se", "intref_prop_se",
                             "intmed_prop_se", "pie_prop_se",
                             "overall_pm_se", "overall_int_se", "overall_pe_se")) {

           mid <- sd(get(stringr::str_replace(effect_se, "_se", "_sim")))

           assign(effect_se, mid)

         }

         # 3 way decomposition effects and se's
         decomp3way <-  c(cde_rr = cde_rr, cde_rr_se = cde_rr_se,
                          pnde_rr = pnde_rr, pnde_rr_se = pnde_rr_se,
                          tnde_rr = tnde_rr, tnde_rr_se = tnde_rr_se,
                          pnie_rr = pnie_rr, pnie_rr_se = pnie_rr_se,
                          tnie_rr = tnie_rr, tnie_rr_se = tnie_rr_se,
                          te_rr = te_rr, te_rr_se = te_rr_se,
                          pm = pm, pm_se = pm_se)

         # 4 way decomposition effects and se's
         decomp4way <-  c(cde_err = cde_err, cde_err_se = cde_err_se, intref_err = intref_err,
                          intref_err_se = intref_err_se, intmed_err = intmed_err,
                          intmed_err_se = intmed_err_se, pie_err = pie_err, pie_err_se = pie_err_se,
                          total_err = total_err, total_err_se = total_err_se,
                          cde_prop = cde_prop, cde_prop_se = cde_prop_se,
                          intref_prop = intref_prop, intref_prop_se = intref_prop_se,
                          intmed_prop = intmed_prop, intmed_prop_se = intmed_prop_se,
                          pie_prop = pie_prop, pie_prop_se = pie_prop_se,
                          overall_pm = overall_pm, overall_pm_se = overall_pm_se,
                          overall_int = overall_int, overall_int_se = overall_int_se,
                          overall_pe = overall_pe, overall_pe_se = overall_pe_se)

         out <- list(decomp3way = decomp3way, decomp4way = decomp4way,
                     sims = data.frame(cde_rr = cde_rr_sim, pnde_rr = pnde_rr_sim, tnde_rr = tnde_rr_sim,
                                       pnie_rr = pnie_rr_sim, tnie_rr = tnie_rr_sim,
                                       cde_err_sim = cde_err_sim, intref_err_sim = intref_err_sim,
                                       intmed_err_sim = intmed_err_sim, pie_err_sim = pie_err_sim))

         }
       }

   return(out)

  }


create_formulas <- function(outcome = NULL, treatment = NULL, mediator = NULL, covariates = c(),
                            interaction = NULL, event = NULL,
                            mreg = c("linear", "logistic"),
                            yreg = c("linear", "logistic", "loglinear", "poisson","quasipoisson",
                                     "negbin", "coxph", "aft_exp", "aft_weibull")) {


  mediator_formula_basic <- paste(mediator, treatment, sep = " ~ ")

  outcome_formula_basic  <- paste(paste(outcome, treatment, sep = " ~ "),
                                  mediator,
                                  sep = " + ")

  if (interaction) {

    outcome_formula_basic <- paste(outcome_formula_basic,
                                   paste(treatment, mediator, sep = " * "),
                                   sep = " + ")
  }

  if (length(covariates) == 0) {

    mediator_formula <- mediator_formula_basic

    outcome_formula  <- outcome_formula_basic

  } else {

    mediator_formula <- paste(mediator_formula_basic,
                              paste(covariates, collapse = " + "),
                              sep = " + ")

    outcome_formula  <- paste(outcome_formula_basic,
                              paste(covariates, collapse = " + "),
                              sep = " + ")
  }

  if (yreg %in% c("coxph","aft_exp","aft_weibull")) {

    l <- strsplit(outcome_formula, split = " ~ ")

    l[[1]][1] <- paste0("Surv(", outcome, ", ", event, ")")

    outcome_formula <- paste(l[[1]][1], l[[1]][2], sep = " ~ ")
  }

  formulas <- list(outcome_formula = outcome_formula,
                   mediator_formula = mediator_formula)

  return(formulas)
}


run_regressions <- function(formulas = NULL, outcome, treatment, mediator, covariates,
                            interaction, event, yreg, mreg, data = NULL) {

  outcome_formula <- formulas$outcome_formula

  mediator_formula<- formulas$mediator_formula

  if (yreg == "linear") {
    outcome_regression  <- lm(outcome_formula, data = data)
  }

  if (yreg == "logistic") {
    outcome_regression  <- glm(outcome_formula, family = binomial(), data = data)
  }

  if (yreg == "loglinear") {
    outcome_regression  <- glm(outcome_formula, family = binomial("log"), data = data)
  }

  if (yreg == "poisson") {
    outcome_regression  <- glm(outcome_formula, family = poisson(), data = data)
  }

  if (yreg == "quasipoisson") {
    outcome_regression  <- glm(outcome_formula, family = quasipoisson(), data = data)
  }

  if (yreg == "negbin") {
    outcome_regression  <- glm.nb(outcome_formula, data = data)
  }

  if (yreg == "coxph") {
    outcome_regression <- survival::coxph(as.formula(outcome_formula), data = data)
  }

  if (yreg == "aft_exp") {
    outcome_regression <- survreg(as.formula(outcome_formula), dist = "exponential", data = data)
  }

  if (yreg == "aft_weibull") {
    outcome_regression <- survreg(as.formula(outcome_formula), dist = "weibull", data = data)
  }

  if (mreg == "linear") {
    mediator_regression <- lm(mediator_formula, data = data)
  }

  if (mreg == "logistic") {
    mediator_regression <- glm(mediator_formula, family = binomial(), data = data)
  }

  regressions <- list(mediator_regression = mediator_regression,
                      outcome_regression = outcome_regression)

  return(regressions)
}


get_coef <- function(regressions = NULL, outcome, treatment, mediator, covariates,
                     interaction, event, mreg, yreg) {

  mediator_regression <- regressions$mediator_regression

  outcome_regression <- regressions$outcome_regression

  ## Store coefficients from regressions
  betas  <- coefficients(mediator_regression)

  thetas <- coefficients(outcome_regression)

  ## Store covariances from regressions
  vcov_betas <- vcov(mediator_regression)

  vcov_thetas <- vcov(outcome_regression)

  if (mreg == "linear")
    variance = summary(mediator_regression)$sigma^2
  else variance = NULL

  if (yreg == "aft_weibull")
    vcov_thetas <- vcov_thetas[-nrow(vcov_thetas), -ncol(vcov_thetas)]

  ## Build block diagonal matrix
  vcov_block <- Matrix::bdiag(vcov_thetas, vcov_betas)

  coef <- list(outcome = outcome, treatment = treatment, mediator = mediator,
               covariates = covariates, interaction = interaction,
               event = event, mediator_reg = mreg, outcome_reg = yreg,betas = betas, thetas=thetas,
               vcov_betas = vcov_betas, vcov_thetas = vcov_thetas, vcov_block = vcov_block,
               variance = variance)

  return(coef)
}


bootstrap_step <- function(data, indices, outcome, treatment, mediator, covariates, vecc,
                                interaction, event, mreg, yreg, m_star, a_star, a) {

  data_boot <- data[indices, ]

  formulas <- create_formulas(outcome = outcome, treatment = treatment,
                              mediator = mediator, covariates = covariates, interaction = interaction,
                              event = event, mreg = mreg, yreg = yreg)

  regressions <- run_regressions(formulas = formulas, outcome = outcome, treatment = treatment,
                                 mediator = mediator, covariates = covariates, interaction = interaction,
                                 event = event, mreg = mreg, yreg = yreg, data = data_boot)

  coef <- get_coef(regressions = regressions, outcome = outcome, treatment = treatment,
                   mediator = mediator, covariates = covariates, interaction = interaction,
                   event = event, mreg = mreg, yreg = yreg)

  thetas <- coef$thetas

  betas <- coef$betas

  variance <- coef$variance

  covariatesTerm <- ifelse(!is.null(vecc), sum(betas[covariates]*t(vecc)), 0)

  interactionTerm <- ifelse(interaction, thetas[paste(treatment, mediator, sep = ":")], 0)

  if (yreg == "linear") {

    if (mreg == "linear") {

      cde <- unname((thetas[treatment] + interactionTerm * m_star) * (a - a_star))

      pnde <- unname((thetas[treatment] + interactionTerm * (betas[1] + betas[treatment] * a_star +
                                                               covariatesTerm))*(a - a_star))

      tnde <- unname((thetas[treatment] + interactionTerm * (betas[1] + betas[treatment] * a +
                                                               covariatesTerm))*(a - a_star))

      pnie <- unname((thetas[mediator] * betas[treatment] +
                        interactionTerm * betas[treatment] * a_star) * (a - a_star))

      tnie <- unname((thetas[mediator] * betas[treatment] +
                        interactionTerm * betas[treatment] * a) * (a - a_star))

    }

    if (mreg == "logistic") {

      cde <- unname((thetas[treatment] + interactionTerm * m_star * (a - a_star)))

      pnde <- unname(thetas[treatment] * (a - a_star) + interactionTerm*(a - a_star) *
                       (exp(betas[1] + betas[treatment] * a_star + covariatesTerm) /
                          (1 + exp(betas[1] + betas[treatment] * a_star + covariatesTerm))))

      tnde <- unname(thetas[treatment] * (a - a_star) + interactionTerm*(a - a_star) *
                       (exp(betas[1] + betas[treatment] * a + covariatesTerm) /
                          (1 + exp(betas[1] + betas[treatment] * a + covariatesTerm))))

      pnie <- unname((thetas[mediator]+interactionTerm*a_star) * (exp(betas[1] +
                                                                        betas[treatment] * a + covariatesTerm) / (1 + exp(betas[1] +
                                                                                                                            betas[treatment] * a + covariatesTerm)) - exp(betas[1] + betas[treatment] * a_star +
                                                                                                                                                                            covariatesTerm) / (1 + exp(betas[1] + betas[treatment] * a_star + covariatesTerm))))

      tnie <- unname((thetas[mediator]+interactionTerm*a) * (exp(betas[1] +
                                                                   betas[treatment] * a + covariatesTerm) / (1 + exp(betas[1] +
                                                                                                                       betas[treatment] * a + covariatesTerm)) - exp(betas[1] + betas[treatment] * a_star +
                                                                                                                                                                       covariatesTerm) / (1 + exp(betas[1] + betas[treatment] * a_star + covariatesTerm))))

      }

    intref <- pnde - cde

    intmed <- tnie - pnie

    pie <- pnie

    te <- tnie + pnde

    pm <- tnie / (pnde + te)

    cde_prop <- cde/te

    intref_prop <- intref/te

    intmed_prop <- intmed/te

    pie_prop <- pie/te

    overall_pm <- (pie+intmed)/te

    overall_int <- (intref+intmed)/te

    overall_pe <- (intref+intmed+pie)/te

    out <- c(cde = cde, pnde = pnde, tnde = tnde, pnie = pnie, tnie = tnie,
             intref = intref, intmed = intmed, pie = pie,
             te = te, pm = pm, cde_prop = cde_prop, intref_prop = intref_prop,
             intmed_prop = intmed_prop, pie_prop = pie_prop,
             overall_pm = overall_pm, overall_int = overall_int, overall_pe = overall_pe)

  }

  if (yreg %in% c("logistic", "loglinear", "poisson", "quasipoisson",
                  "negbin", "coxph", "aft_exp", "aft_weibull")) {

    if (mreg == "linear") {

      cde_rr <- unname(exp((thetas[treatment] + interactionTerm * m_star) * (a - a_star)))

      pnde_rr <- unname(exp((thetas[treatment] + interactionTerm * (betas[1] + betas[treatment] * a_star +
                                                                   covariatesTerm + thetas[mediator]  * variance)) * (a - a_star) +
                           0.5 * interactionTerm ^ 2 * variance * (a^2 - a_star ^ 2)))

      tnde_rr <- unname(exp((thetas[treatment] + interactionTerm * (betas[1] + betas[treatment] * a +
                                                                   covariatesTerm + thetas[mediator]  * variance)) * (a - a_star) +
                           0.5 * interactionTerm ^ 2 * variance * (a^2 - a_star ^ 2)))

      pnie_rr <- unname(exp((thetas[mediator] * betas[treatment] +
                            interactionTerm * betas[treatment] * a_star) * (a - a_star)))

      tnie_rr <- unname(exp((thetas[mediator] * betas[treatment] +
                            interactionTerm * betas[treatment] * a) * (a - a_star)))

      cde_err <- unname(exp(thetas[treatment]*(a-a_star)+thetas[mediator]*m_star+
                              interactionTerm*a*m_star- (thetas[mediator]+interactionTerm*a_star)*
                               (betas['(Intercept)']+betas[treatment]*a_star+sum(betas[covariates]*t(vecc)))-
                               0.5*(thetas[mediator]+interactionTerm*a_star)^2*coef$variance)-
                           exp(thetas[mediator]*m_star+interactionTerm*a_star*m_star-
                                 (thetas[mediator]+interactionTerm*a_star)*(betas['(Intercept)']+
                                                                                betas[treatment]*a_star+sum(betas[covariates]*t(vecc)))-
                                 0.5*(thetas[mediator]+interactionTerm*a_star)^2*coef$variance))

    }

    if (mreg == "logistic") {

      cde_rr <- unname(exp((thetas[treatment] + interactionTerm*m_star) * (a - a_star)))

      pnde_rr <- unname((exp(thetas[treatment] * (a - a_star)) * (1 + exp(thetas[mediator] +
                                                                         interactionTerm * a + betas[1] + betas[treatment] * a_star +
                                                                         covariatesTerm))) /(1 + exp(thetas[mediator] + interactionTerm * a_star +
                                                                                                       betas[1] + betas[treatment] * a_star + covariatesTerm)))

      tnde_rr <- unname((exp(thetas[treatment] * (a - a_star)) * (1 + exp(thetas[mediator] +
                                                                         interactionTerm * a + betas[1] + betas[treatment] * a + covariatesTerm))) /
                       (1 + exp(thetas[mediator] + interactionTerm * a_star +  betas[1] +
                                  betas[treatment] * a + covariatesTerm)))

      pnie_rr <- unname(((1 + exp(betas[1] + betas[treatment] * a_star + covariatesTerm)) *
                        (1 + exp(thetas[mediator] + interactionTerm * a_star + betas[1] +
                                   betas[treatment] * a + covariatesTerm))) /
                       ((1 + exp(betas[1] + betas[treatment] * a + covariatesTerm)) *
                          (1+ exp(thetas[mediator] + interactionTerm * a_star + betas[1] +
                                    betas[treatment] * a_star + covariatesTerm))))

      tnie_rr <- unname(((1 + exp(betas[1] + betas[treatment] * a_star + covariatesTerm)) *
                        (1 + exp(thetas[mediator] + interactionTerm * a + betas[1] + betas[treatment] * a +
                                   covariatesTerm))) / ((1 + exp(betas[1] + betas[treatment] * a + covariatesTerm)) *
                                                          (1 + exp(thetas[mediator] + interactionTerm * a + betas[1] + betas[treatment] * a_star +
                                                                     covariatesTerm))))

      cde_err <- unname((exp(thetas[treatment]*(a-a_star)+thetas[mediator]*m_star+
                               interactionTerm*a*m_star)*(1+exp(betas['(Intercept)']+
                  betas[treatment]*a_star+sum(betas[covariates]*t(vecc))))/(1+exp(betas['(Intercept)']+
                  betas[treatment]*a_star+sum(betas[covariates]*t(vecc))+thetas[mediator]+
                    interactionTerm*a_star))-exp(thetas[mediator]*m_star+
                                                   interactionTerm*a_star*m_star)*(1+exp(betas['(Intercept)']+
                  betas[treatment]*a_star+sum(betas[covariates]*t(vecc))))/(1+exp(betas['(Intercept)']+
                  betas[treatment]*a_star+sum(betas[covariates]*t(vecc))+thetas[mediator]+
                    interactionTerm*a_star))))

    }

    intref_err <- pnde_rr - 1 - cde_err

    intmed_err <- tnie_rr * pnde_rr - pnde_rr - pnie_rr + 1

    pie_err <- pnie_rr - 1

    te_rr <- tnie_rr * pnde_rr

    pm <- (pnde_rr * (tnie_rr - 1)) / (pnde_rr * tnie_rr - 1)

    total_err <- te_rr - 1

    cde_err_prop <- cde_err/total_err

    intmed_err_prop <- intmed_err/total_err

    intref_err_prop <- intref_err/total_err

    pie_err_prop <- pie_err/total_err

    overall_pm <- (pie_err+intmed_err)/total_err

    overall_int <- (intref_err+intmed_err)/total_err

    overall_pe <- (intref_err+intmed_err+pie_err)/total_err

    out <- c(cde_rr = cde_rr, pnde_rr = pnde_rr, tnde_rr = tnde_rr, pnie_rr = pnie_rr,
             tnie_rr = tnie_rr, te_rr = te_rr, pm = pm,
             cde_err = cde_err, intref_err = intref_err,
             intmed_err = intmed_err, pie_err = pie_err, total_err = total_err,
             cde_err_prop = cde_err_prop, intref_err_prop = intref_err_prop,
             intmed_err_prop = intmed_err_prop, pie_err_prop = pie_err_prop,
             overall_pm = overall_pm, overall_int = overall_int,
             overall_pe = overall_pe)


  }

  return(out)

}


z_p <- function(s, n, nway, yreg) {

  z <- NULL

  pval <- NULL

  if (nway == 3) {

    if (yreg == "linear") {

      z <- s$estimate / s$std.error

      pval <- 2 * pt(-abs(z), n - 1)

      } else {

      z[1:6] <- (s$estimate[1:6]-1) / s$std.error[1:6]

      z[7] <- s$estimate[7] / s$std.error[7]

      pval <- 2 * pt(-abs(z), n - 1)

      }
    } else if (nway == 4) {

        z <- s$estimate / s$std.error

        pval <- 2 * pt(-abs(z), n - 1)

        }

  return(data.frame(z = z, pval = pval))

}

format_row_boot <- function(boot.out, index = 1, conf = 0.95) {
  ci <- boot::boot.ci(boot.out, index = index, type = "perc", conf = conf)
  d <- cbind(as.data.frame(t(tail(ci$percent[1, ], 2))))
  return(d)
}


format_df_boot <- function(boot.out, conf = 0.95, n, nway, yreg) {

  d_all <- NULL

  for (i in 1:length(boot.out$t0)) {

    d <- format_row_boot(boot.out, i, conf = conf)

    d_all <- rbind(d_all, d)

  }

  estimate <- boot.out$t0

  bias <- apply(boot.out$t, 2, mean) - estimate

  std.error <- apply(boot.out$t, 2, sd)

  d_all <- cbind(estimate, bias, std.error, d_all)

  rownames(d_all) <- names(boot.out$t0)

  label_CI <- paste0(round(conf * 100, 2), c("% CIL", "% CIU"))

  colnames(d_all) <- c("estimate", "bias", "std.error", label_CI[1], label_CI[2])


  return(cbind(d_all, z_p(d_all, n = n, nway = nway, yreg = yreg)))

}

format_df_delta <- function(delta.out, conf = 0.95, n, nway, yreg) {

  d_all <- data.frame(matrix(NA, length(delta.out)/2, 4))

  alpha <- (1 - conf) / 2

  z <- qnorm(1 - alpha)

  label_CI <- paste0(round(conf * 100, 2), c("% CIL", "% CIU"))

  colnames(d_all) <- c("estimate", "std.error", label_CI[1], label_CI[2])

  rownames(d_all) <- names(delta.out)[2*(1:(length(delta.out)/2))-1]

  for (name in rownames(d_all)) {

    d_all[name, ] <- c(unname(delta.out[name]), unname(delta.out[stringr::str_c(name,"se_delta",sep="_")]),
                       unname(delta.out[name]-z*delta.out[stringr::str_c(name,"se_delta",sep="_")]),
                       unname(delta.out[name]+z*delta.out[stringr::str_c(name,"se_delta",sep="_")]))

  }

  return(cbind(d_all, z_p(d_all, n = n, nway = nway, yreg = yreg)))

}

print.delta_out <- function(delta_out, digits = 2) {

  printCoefmat(delta_out, digits = digits, has.Pvalue = TRUE)

}

print.boot_out <- function(boot_out, digits = 2) {

  printCoefmat(boot_out, digits = digits, has.Pvalue = TRUE)

}


df <- read.csv("int_data_10000.txt", sep = " ")
```

# Case 1: Continuous Outcome and Continuous Mediator

## Data simulation

### Simulation Procedures

1. Simulate the exposure variable A from Binom(1,P(A=1)).

2. Simulate the covariate C from $N(\mu_C,\sigma_C^2)$.

3. Simulate the mediator M from $N((\beta_0+\beta_1*A+\beta_2*C),\sigma_M^2)$.

4. When there exists interaction between the exposure and the mediator, simulate the outcome Y from $N(\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4C,\sigma_Y^2)$; When there doesn't exist interaction between the exposure and the mediator, simulate the outcome Y from $N(\theta_0+\theta_1A+\theta_2M+\theta_4C,\sigma_Y^2)$.

### True Parameters

\begin{center}
\captionof{table}{True Model Parameters for Data Simulation}
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l|l|l}
\hline
Sample Size & $\theta_0$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\beta_0$ & $\beta_1$ & $\beta_2$ & P(A=1) & $\mu_C$ & $\sigma_C$ & $\sigma_M$ & $\sigma_Y$\\ \hline
10000&-3.2&0.1&0.1&0.1&0.1&-0.25&0.3&0.2&0.4&1&1&1&1 \\ \hline
\end{tabular}
\end{center}

### True Models and True Causal Effects

True model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2c$$

True model for the outcome:
$$E[Y|a,m^*,c]=\theta_0+\theta_1a+\theta_2m^*+\theta_3am+\theta_4c$$

True Causal Effects with $m^*=0$, $a^*=0$ and $a=1$:

\begin{center}
\captionof{table}{True Causal Effects from 3 Way Decomposition for Cont Y and Cont M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
cde & pnde & tnde & pnie & tnie & total effect & proportion mediated \\ \hline
0.100&0.095&0.125&0.030&0.060&0.155&0.240 \\ \hline
\end{tabular}
\end{center}

\begin{center}
\captionof{table}{True Causal Effects from 4 Way Decomposition for Cont Y and Cont M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
cde & intref & intmed & pie & total effect & cde prop & intref prop \\ \hline
0.100&-0.005&0.030&0.030&0.155&0.645&-0.032 \\ \hline
intmed prop & pie prop&overall pm&overall int&overall pe\\ \hline
0.194&0.194&0.387&0.161&0.355 \\ \hline
\end{tabular}
\end{center}

### Estimated Causal Effects and Standard Errors from 3 Different Estimation Method

#### Delta Method

```{r}
causal_mediation(data = df, method="delta", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "linear")
```

#### Bootstrapping

```{r}
causal_mediation(data = df, method="bootstrap", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "linear")
```

#### Simulation-based Approach

```{r}
causal_mediation(data = df, method="simulation", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "linear")[1:2]
```


# Case 2: Continuous Outcome and Binary Mediator

## Data simulation

### Simulation Procedures

1. Simulate the exposure variable A from Binom(1,P(A=1)).

2. Simulate the covariate C from $N(\mu_C,\sigma_C^2)$.

3. Simulate the mediator M from $Binom(1,expit(\beta_0+\beta_1*A+\beta_2*C))$.

4. When there exists interaction between the exposure and the mediator, simulate the outcome Y from $N(\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4C,\sigma_Y^2)$; When there doesn't exist interaction between the exposure and the mediator, simulate the outcome Y from $N(\theta_0+\theta_1A+\theta_2M+\theta_4C,\sigma_Y^2)$.

### True Parameters

\begin{center}
\captionof{table}{True Model Parameters for Data Simulation}
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l|l}
\hline
Sample Size & $\theta_0$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\beta_0$ & $\beta_1$ & $\beta_2$ & P(A=1) & $\mu_C$ & $\sigma_C$ & $\sigma_Y$\\ \hline
10000&-3.2&0.1&0.1&0.1&0.1&-0.25&0.3&0.2&0.4&1&1&1 \\ \hline
\end{tabular}
\end{center}

### True Models and True Causal Effects

True model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2c$$

True model for the outcome:
$$E[Y|a,m^*,c]=\theta_0+\theta_1a+\theta_2m^*+\theta_3am+\theta_4c$$

True Causal Effects with $m^*=0$, $a^*=0$ and $a=1$:

\begin{center}
\captionof{table}{True Causal Effects from 3 Way Decomposition for Cont Y and Bin M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
cde & pnde & tnde & pnie & tnie & total effect & proportion mediated \\ \hline
0.100&0.149&0.156&0.007&0.015&0.107&0.058 \\ \hline
\end{tabular}
\end{center}

\begin{center}
\captionof{table}{True Causal Effects from 4 Way Decomposition for Cont Y and Bin M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
cde & intref & intmed & pie & total effect & cde prop & intref prop \\ \hline
0.100&0.049&0.008&0.007&0.107&0.935&0.052 \\ \hline
intmed prop & pie prop&overall pm&overall int&overall pe\\ \hline
0.075&0.065&0.140&0.533&0.598 \\ \hline
\end{tabular}
\end{center}

### Estimated Causal Effects and Standard Errors from 3 Different Estimation Method

#### Delta Method

```{r}
causal_mediation(data = df, method="delta", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "logistic")
```

#### Bootstrapping

```{r}
causal_mediation(data = df, method="bootstrap", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "logistic")
```

#### Simulation-based Approach

```{r}
causal_mediation(data = df, method="simulation", outcome = "Y_cont_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "linear", mreg = "logistic")[1:2]
```


# Case 3: Binary Outcome and Continuous Mediator

## Data simulation

### Simulation Procedures

1. Simulate the exposure variable A from Binom(1,P(A=1)).

2. Simulate the covariate C from $N(\mu_C,\sigma_C^2)$.

3. Simulate the mediator M from $N((\beta_0+\beta_1*A+\beta_2*C),\sigma_M^2)$.

4. When there exists interaction between the exposure and the mediator, simulate the outcome Y from $Binom(1, expit(\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4C,\sigma_Y^2))$; When there doesn't exist interaction between the exposure and the mediator, simulate the outcome Y from $Binom(1, expit(\theta_0+\theta_1A+\theta_2M+\theta_4C,\sigma_Y^2))$.

### True Parameters

\begin{center}
\captionof{table}{True Model Parameters for Data Simulation}
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l|l}
\hline
Sample Size & $\theta_0$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\beta_0$ & $\beta_1$ & $\beta_2$ & P(A=1) & $\mu_C$ & $\sigma_C$ & $\sigma_M$\\ \hline
10000&-3.2&0.1&0.1&0.1&0.1&-0.25&0.3&0.2&0.4&1&1&1 \\ \hline
\end{tabular}
\end{center}

### True Models and True Causal Effects

True model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2c$$

True model for the outcome:
$$logitE[Y|a,m^*,c]=\theta_0+\theta_1a+\theta_2m^*+\theta_3am+\theta_4c$$

True Causal Effects with $m^*=0$, $a^*=0$ and $a=1$:

\begin{center}
\captionof{table}{True Causal Effects from 3 Way Decomposition for Bin Y and Cont M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
RRcde & RRpnde & RRtnde & RRpnie & RRtnie & RRte & proportion mediated \\ \hline
1.105& 1.116&1.150&1.030&1.062&1.185&0.373 \\ \hline
\end{tabular}
\end{center}

\begin{center}
\captionof{table}{True Causal Effects from 4 Way Decomposition for Bin Y and Cont M}
\begin{tabular}{l|l|l|l|l|l}
\hline
ERRcde & ERRintref & ERRintmed & ERRpie & ERRte & ERRcde prop \\ \hline
0.105&0.011&0.039&0.030&0.185&0.568 \\ \hline
ERRintref prop&ERRintmed prop&ERRpie prop&overall pm&overall int&overall pe \\ \hline
0.059&0.211&0.162&0.373&0.270&0.432 \\ \hline
\end{tabular}
\end{center}

### Estimated Causal Effects and Standard Errors from 3 Different Estimation Method

#### Delta Method

```{r}
causal_mediation(data = df, method="delta", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "linear")
```

#### Bootstrapping

```{r}
causal_mediation(data = df, method="bootstrap", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "linear")
```

#### Simulation-based Approach

```{r}
causal_mediation(data = df, method="simulation", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_cont', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "linear")[1:2]
```


# Case 4: Binary Outcome and Binary Mediator

## Data simulation

### Simulation Procedures

1. Simulate the exposure variable A from Binom(1,P(A=1)).

2. Simulate the covariate C from $N(\mu_C,\sigma_C^2)$.

3. Simulate the mediator M from $Binom(1,expit(\beta_0+\beta_1*A+\beta_2*C))$.

4. When there exists interaction between the exposure and the mediator, simulate the outcome Y from $Binom(1, expit(\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4C,\sigma_Y^2))$; When there doesn't exist interaction between the exposure and the mediator, simulate the outcome Y from $Binom(1, expit(\theta_0+\theta_1A+\theta_2M+\theta_4C,\sigma_Y^2))$.

### True Parameters

\begin{center}
\captionof{table}{True Model Parameters for Data Simulation}
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l}
\hline
Sample Size & $\theta_0$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\beta_0$ & $\beta_1$ & $\beta_2$ & P(A=1) & $\mu_C$ & $\sigma_C$ \\ \hline
10000&-3.2&0.1&0.1&0.1&0.1&-0.25&0.3&0.2&0.4&1&1 \\ \hline
\end{tabular}
\end{center}

### True Models and True Causal Effects

True model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2c$$

True model for the outcome:
$$logitE[Y|a,m^*,c]=\theta_0+\theta_1a+\theta_2m^*+\theta_3am+\theta_4c$$

True Causal Effects with $m^*=0$, $a^*=0$ and $a=1$:

\begin{center}
\captionof{table}{True Causal Effects from 3 Way Decomposition for Bin Y and Bin M}
\begin{tabular}{l|l|l|l|l|l|l}
\hline
RRcde & RRpnde & RRtnde & RRpnie & RRtnie & RRte & proportion mediated \\ \hline
1.105& 1.165&1.173&1.007&1.015&1.182&0.095 \\ \hline
\end{tabular}
\end{center}

\begin{center}
\captionof{table}{True Causal Effects from 4 Way Decomposition for Bin Y and Bin M}
\begin{tabular}{l|l|l|l|l|l}
\hline
ERRcde & ERRintref & ERRintmed & ERRpie & ERRte & ERRcde prop \\ \hline
0.100&0.065&0.010&0.007&0.182&0.549 \\ \hline
ERRintref prop&ERRintmed prop&ERRpie prop&overall pm&overall int&overall pe \\ \hline
0.357&0.055&0.038&0.093&0.412&0.451 \\ \hline
\end{tabular}
\end{center}

### Estimated Causal Effects and Standard Errors from 3 Different Estimation Method

#### Delta Method

```{r}
causal_mediation(data = df, method="delta", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "logistic")
```

#### Bootstrapping

```{r}
causal_mediation(data = df, method="bootstrap", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "logistic")
```

#### Simulation-based Approach

```{r}
causal_mediation(data = df, method="simulation", outcome = "Y_bin_int", treatment = 'A',
                 mediator = 'M_bin', covariates = "C", interaction = TRUE,
                 yreg = "logistic", mreg = "logistic")[1:2]
```
