---
title: "Summary"
author: "Baoyi Shi"
output: 
  pdf_document:
    number_sections: true
header-includes:
- \usepackage{caption}
---

# Supported Data Types and Functionalities

rb: Regression-based Approach

wb: Weighting-based Approach

msm: Marginal Structural Model

iorw: Inverse Odds Ratio Weighting Approach

ne: Natural Effect Model

g-formula: G-formula Approach

\begin{center}
\captionof{table}{Supported Data Types and Functionalities For Single-mediator Cases}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline

&rb & wb & msm & iorw & ne & g-formula \\ \hline
Linear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Loglinear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Poisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Quasipoisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
NegBin Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Coxph Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Exp Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Weibull Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Linear M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Categorical M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Any type M & $\times$& $\surd$& $\times$ & $\surd$& $\surd$& $\times$ \\ \hline
User-defined Y/M Models & $\surd$& $\surd$& $\surd$ & $\surd$& $\times$& $\surd$ \\ \hline
Pre-exposure Confounding & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Post-exposure Confounding & $\times$ & $\times$& $\surd$& $\times$& $\times$& $\surd$ \\ \hline
2-way Decomposition & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
4-way Decomposition & $\surd$& $\surd$& $\surd$& $\times$& $\times$& $\surd$ \\ \hline
Estimation: Closed-form Parameter Function & $\surd$& $\times$& $\surd$& $\surd$& $\times$& $\times$ \\ \hline
Estimation: Direct Imputation & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Inference: Delta Method & $\surd$& $\times$& $\surd$& $\surd$& $\times$& $\times$ \\ \hline
Inference: Bootstrapping & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
\end{tabular}
\end{center}

\begin{center}
\captionof{table}{Supported Data Types and Functionalities For Multiple-mediator Cases}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline

&rb & wb & msm & iorw & ne & g-formula \\ \hline
Linear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Loglinear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Poisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Quasipoisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
NegBin Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Coxph Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Exp Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Weibull Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Linear M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Categorical M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Any type M & $\times$& $\surd$& $\times$ & $\surd$& $\surd$& $\times$ \\ \hline
User-defined Y/M Models & $\surd$& $\surd$& $\surd$ & $\surd$& $\times$& $\surd$ \\ \hline
Pre-exposure Confounding & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Post-exposure Confounding & $\times$ & $\times$& $\surd$& $\times$& $\times$& $\surd$ \\ \hline
2-way Decomposition & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
4-way Decomposition & $\surd$& $\surd$& $\surd$& $\times$& $\times$& $\surd$ \\ \hline
Estimation: Closed-form Parameter Function & $\times$& $\times$& $\times$& $\surd$& $\times$& $\times$ \\ \hline
Estimation: Direct Imputation & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Inference: Delta Method & $\times$& $\times$& $\times$& $\surd$& $\times$& $\times$ \\ \hline
Inference: Bootstrapping & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
\end{tabular}
\end{center}

# Estimation and Inference

## Estimation Method

**Closed-form Parameters Function**

Effect estimates are calculated using regression parameters. 

**Direct Conterfacturals imputation**

1. For subject i, simulate $M_{a,i}$ by the predicted values of mediator regression models under $A=a,C=C_i$ and simulate $M_{a^*,i}$ by the predicted values of mediator regression models under $A=a^*,C=C_i$.

2. For subject i, simulate $Y_{aMa,i}$ by the predicted value of the outcome regression model under $A=a,M=M_{a,i},C=C_i$, simulate $Y_{aMa^*,i}$ by the predicted value of the outcome regression model under $A=a,M=M_{a^*,i},C=C_i$, simulate $Y_{a^*Ma,i}$ by the predicted value of the outcome regression model under $A=a^*,M=M_{a,i},C=C_i$ and simulate $Y_{a^*Ma^*,i}$ by the predicted value of the outcome regression model under $A=a^*,M=M_{a^*,i},C=C_i$. 

3. Estimate $E[Y_{aMa}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa^*}]$, and $E[Y_{a^*Ma}]$ by $\frac{\sum_{i=1}^NY_{aMa,i}}{n}$, $\frac{\sum_{i=1}^NY_{a^*Ma^*,i}}{n}$, $\frac{\sum_{i=1}^NY_{aMa^*,i}}{n}$ and $\frac{\sum_{i=1}^NY_{a^*Ma,i}}{n}$ respectively.

## Inference Method

**Delta Method**

Standard errors of effects are estimated using the standard errors of regression parameters and delta method based on the closed-form parameters function.

**Bootstrapping**

Bootstrap the data, refit the regression models, and calculate a bootstrap estimate. Repeat the bootstrapping K times and calculate the standard error of these K boostrap estimates for each estimand, which is the estimated standard error of this estimand.

# Pre-treatment Confounding

## DAG

```{r echo=F,fig.width=6,fig.height=3}
g1 <- dagitty::dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
}")

dagitty::coordinates(g1) <- list(x=c(A=0,Y=4,M=2,C=0.2),y=c(A=0,Y=0,M=-0.8,C=-1))

plot(g1)
```

In the DAG, A denotes the treatment, Y denotes the outcome, M denotes a set of mediators, and C denotes a set of pre-treatment covariates.

## Estimand

**2-way decomposition in additive scale**

$CDE=E[Y_{am}-Y_{a^*m}|C]$

$PNDE=E[Y_{aM_a^*}-Y_{a^*M_a^*}|C]$

$TNDE=E[Y_{aM_a}-Y_{a^*M_a}|C]$

$PNIE=E[Y_{a^*M_a}-Y_{a^*M_a^*}|C]$

$TNIE=E[Y_{aM_a}-Y_{aM_a^*}|C]$

$TE=PNDE+TNIE$


**2-way decomposition in RR scale**

$rr\_CDE=E[Y_{am}|C]/E[Y_{a^*m}|C]$

$rr\_PNDE=E[Y_{aM_a^*}|C]/E[Y_{a^*M_a^*}|C]$

$rr\_TNDE=E[Y_{aM_a}|C]/E[Y_{a^*M_a}|C]$

$rr\_PNIE=E[Y_{a^*M_a}|C]/E[Y_{a^*M_a^*}|C]$

$rr\_TNIE=E[Y_{aM_a}|C]/E[Y_{aM_a^*}|C]$

$rr\_TE=rr\_PNDE\times rr\_TNIE$


**4-way decomposition in additive scale**

$CDE$ is defined above

$INT_{ref}=PNDE-CDE$

$INT_{med}=TNIE-PNIE$

$PIE=PNIE$


**4-way decomposition in RR scale**

$err\_CDE=(E[Y_{am}-Y_{a^*m}|C])/E[Y_{a^*M_a^*}|C]$

$err\_INT_{ref}=rr\_PNDE-1-err\_CDE$

$err\_INT_{med}=rr\_TNIE*rr\_PNDE-rr\_PNDE-rr\_PNIE+1$

$err\_PIE=rr\_PNIE-1$

$err\_TE=err\_CDE+err\_INT_{ref}+err\_INT_{med}+err\_PIE=rr\_{TE}-1$

## Estimation Approaches Can be Used

### Regression-based Approach

Reference:

https://www.ncbi.nlm.nih.gov/pubmed/23379553

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/

#### Estimator

**1. Continuous Outcome and Continuous Mediator**

Fit a simple linear regression model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2'c$$
Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$CDE=(\theta_1+\theta_3m)(a-a^*)$

$PNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c)\}(a-a^*)$

$TNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c)\}(a-a^*)$

$PNIE=(\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*)$

$TNIE=(\theta_2\beta_1+\theta_3\beta_1a)(a-a^*)$

&nbsp; 

**2. Continuous Outcome and Binary Mediator**

Fit a logistic regression model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2'c$$
Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$CDE=(\theta_1+\theta_3m)(a-a^*)$

$PNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)}\}(a-a^*)$

$TNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}\}(a-a^*)$

$PNIE=(\theta_2+\theta_3a^*)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$

$TNIE=(\theta_2+\theta_3a)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$

&nbsp; 

**3. Continuous Outcome and Categorical Mediator**

Fit a multinomial logistic regression model for the mediator:

$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\beta_{1j}a+\beta_{2j}'c, j=1,2,...,l$$, where l is the number of levels of M and $E[M=0|a,c]=\frac{1}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}$.

Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\sum_{j=1}^l\theta_{2j}I\{m=j\}+a\sum_{j=1}^l\theta_{3j}I\{m=j\}+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$CDE=(\theta_1+\sum_{j=1}^l\theta_{3j}I\{m=j\})(a-a^*)$

$PNDE=\{\theta_1+\frac{\sum_{j=1}^l\theta_{3j}exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}\}(a-a^*)$

$TNDE=\{\theta_1+\frac{\sum_{j=1}^l\theta_{3j}exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}\}(a-a^*)$

$PNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

$TNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

&nbsp; 

**4. Binary Outcome and Continuous Mediator**

Fit a simple linear regression model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2'c$$

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$OR^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$

$OR^{PNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$

$OR^{TNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$

$OR^{PNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*))$

$OR^{TNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a)(a-a^*))$

$comp^{CDE}=(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))exp(\theta_2m-(\theta_2+\theta_3a^*)(\beta_0+\beta_1a^*+\beta_2'c)-0.5(\theta_2+\theta_3a^*)^2\sigma^2)$

&nbsp; 

**5. Binary Outcome and Binary Mediator**

Fit a logistic regression model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2'c$$

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$OR^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$

$OR^{PNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$OR^{TNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}$

$OR^{PNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$OR^{TNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$comp^{CDE}=\frac{exp(\theta_2m)(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))(1+exp(\beta_0+\beta_1a^*+\beta_2'c))}{1+exp(\beta_0+\beta_1a^*+\beta_2'c+\theta_2+\theta_3a^*)}$

&nbsp; 

**6. Binary Outcome and Categorical Mediator**

Fit a multinomial logistic regression model for the mediator:
$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\beta_{1j}a+\beta_{2j}'c, j=1,2,...,l$$, where l is the number of levels of M and $E[M=0|a,c]=\frac{1}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}$.

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c$$

Closed-form parameter function estimators for the causal effects:

$OR^{CDE}=exp((\theta_1+\sum_{j=1}^l\theta_{3j}I\{m=j\})(a-a^*))$

$OR^{PNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$OR^{TNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}$

$OR^{PNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$OR^{TNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$comp^{CDE}=\frac{exp(\sum_{j=1}^l\theta_{2j}I\{m=j\})(exp(\theta_1(a-a^*)+a\sum_{j=1}^l\theta_{3j}I\{m=j\})-exp(a^*\sum_{j=1}^l\theta_{3j}I\{m=j\}))(1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c))}{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

#### Estimation Procedures

1. Fit a regression model for Y on A, M and C.

2. Fit a regression model for each mediator in M on A and C.

3. Estimate the point estimates of causal effects using closed-form parameters functions or direct Conterfacturals imputation.

4. Estimate the standard errors of causal effects using delta method or bootstrapping.

### Weighting-based Approach

Reference:

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/

**Procedures**

1. Fit a regression model for Y on A, M, C.

2. For $E[Y_{a^*Ma^*}]$, estimate it by taking a weighted average of the subjects with $A=a^*$ and each subject i is given a weight $\frac{P(A=a^*)}{P(A=a^*|c_i)}$.

3. For $E[Y_{aMa}]$, estimate it by taking a weighted average of the subjects with $A=a$ and each subject i is given a weight $\frac{P(A=a)}{P(A=a|c_i)}$.

4. For $E[Y_{aMa^*}]$, estimate it by taking a weighted average of the predicted Y values for subjects with $A=a^*$ and each subject i is given a weight $\frac{P(A=a^*)}{P(A=a^*|c_i)}$.

5. For $E[Y_{a^*Ma}]$, estimate it by taking a weighted average of the predicted Y values for subjects with $A=a$ and each subject i is given a weight $\frac{P(A=a)}{P(A=a|c_i)}$.

6. Use bootsrapping to estimate the standard error for each estimand.

### Inverse Odds Ratio Weighting Approach

Reference:

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3954805/

https://www.ncbi.nlm.nih.gov/pubmed/25693776

**Procedures**

1. Fit a model for A given M and C

2. Calculate the stablized weights for each subject, $w_i=\frac{f_{A|M,C}(A=0|M_i,C_i)}{f_{A|M,C}(A=A_i|M_i,C_i)}$.

3. Estimate the direct effect by a weighted regression model of Y on A and C using the weights calculated in 2. The estimated direct effect is the coefficient of A in this regression model.

4. Estimate the total effect by a regression model of Y on A and C. The estimated total effect is the coefficient of A in this regression model.

5. Calculate the indirect effect by subtracting the direct effect from the total effect.

### Natural Effect Model

Incorporate the Medflex package.

### Other Approaches

Approaches talked about later can also be used.

# Post-treatment Confounding

## DAG

```{r echo=F,fig.width=6,fig.height=4}
g1 <- dagitty::dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
    A -> L
    L -> Y
    L -> M
    C -> L
}")

dagitty::coordinates(g1) <- list(x=c(A=0,Y=4,M=2,C=0.2,L=2),y=c(A=0,Y=0,M=-0.8,C=-1,L=0.5))

plot(g1)
```

In the DAG, A denotes the treatment, Y denotes the outcome, M denotes a set of mediators and $M=(M_1, M_2,...,M_n)$, C denotes a set of pre-treatment covariates and L denotes a set of post-treatment covariates and $L=(L_1, L_2,...,L_p)$.

## Estimand

$rNDE=E(Y_{aGa^*})-E(Y_{a^*Ga^*})=\sum_{l,m,c}\{E[Y|a,l,m,c]P(l|a,c)-E[Y|a^*,l,m,c]P(l|a^*,c)\}P(m|a^*,c)P(c)$

$rNIE=E(Y_{aGa})-E(Y_{aGa^*})=\sum_{l,m,c}\{E[Y|a,l,m,c]P(l|a,c)\{P(m|a,c)-P(m|a^*,c)\}P(c)$

$rTE=rNDE+rNIE$

## Estimation Approaches Can be Used

### Marginal Structual Model

Reference:

https://journals.lww.com/epidem/fulltext/2009/01000/marginal_structural_models_for_the_estimation_of.6.aspx

**Procedures**

For rCDE:

1. Fit a weighted regression model for Y on A and M where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}\frac{P(M_1=m_{1i}|A=a_i)}{P(M_1=m_{1i}|A=a_i,C=c_i,L=l_i)}...\frac{P(M_n=m_{ni}|A=a_i)}{P(M_n=m_{ni}|A=a_i,C=c_i,L=l_i)}$.

2. Get the point estimate using the closed-form parameter function or direct imputation.

3. Get the standard error using the delta method or bootstrapping.

For rNDE and rNIE:

1. Fit a weighted regression model for Y on A, M and C where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}\frac{P(M_1=m_{1i}|A=a_i)}{P(M_1=m_{1i}|A=a_i,C=c_i,L=l_i)}...\frac{P(M_n=m_{ni}|A=a_i)}{P(M_n=m_{ni}|A=a_i,C=c_i,L=l_i)}$.

2. Fit a weighted regression model for each $M_1, M_2,...,M_n$ on A and C where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}$.

3. For single-mediator cases, get the point estimate using the closed-form parameter function or direct imputation; for multiple-mediator cases, get the point estimate using direct imputation.

4. For single-mediator cases, get the point estimate using the delta method or bootstrapping; for multiple-mediator cases, get the point estimate using bootstrapping.

### G-formula Approach

Reference: 

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5285457/

**Estimation Algorithm for $E(Y_{a1Ga2})$**

1. Bootstrap the dataset.

2. Fit models for $E(L_1|A,C)$, $E(L_2|A,C)$,..., $E(L_p|A,C)$ using the bootstrapped dataset. Then, for each subject i, simulate the value of $L^*_{pi}$ under A=a1 by the predicted value of $L_{pi}|A=a_1,C=C_i$.

3. Fit models for $E(M_1|A,C,L)$, $E(M_2|A,C,L)$,...,$E(M_n|A,C,L)$ using the bootstrapped dataset. Then, for each subject i, simulate the value of $M^*_{ni}$ under A=a2 by the predicted value of $M_{ni}|A=a2,C=C_i,L=L_i$.

4. Fit models for E(Y|A,L,M,C) using the bootstrapped dataset. Then, for each subject i, simulate the potential outcome $Y_i^*$ by the predicted value of $Y|A=a1,L_p=L^*_{pi},...,L_2=L^*_{2i},L_1=L^*_{1i},M_n=M^*_{ni},...,M_2=M^*_{2i},M_1=M^*_{1i},C=C_i,$.

5. Calculate mean value of $Y^*$, i.e, $\sum_{i=1}^NY_i^*$

6. Repeat 1-5 K times and estimate $E(Y_{a1Ga2})$ as $\sum_{k=1}^K\sum_{i=1}^NY_{ik}^*$.
