---
title: "Summary"
author: "Baoyi Shi"
output: 
  pdf_document:
    number_sections: true
header-includes:
- \usepackage{caption}
- \usepackage{threeparttable}
---

# Supported Data Types and Functionalities

\begin{center}
\begin{threeparttable}
\captionof{table}{Supported Data Types and Functionalities}
\begin{tabular}{|l|l|l|l|l|l|l|}

\hline

&rb & wb & msm & iorw & ne & g-formula \\ \hline
Linear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Loglinear Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Poisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Quasipoisson Y & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
NegBin Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Coxph Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Exp Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
AFT Weibull Y & $\surd$& $\surd$& $\surd$& $\surd$& $\times$& $\surd$ \\ \hline
Linear M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Logistic M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Categorical M & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Any type M & $\oslash$& $\surd$& $\oslash$ & $\surd$& $\surd$& $\oslash$ \\ \hline
User-defined Y/M Models & $\surd$& $\surd$& $\surd$ & $\surd$& $\times$& $\surd$ \\ \hline
Continuous A & $\surd$& $\oslash$& $\oslash$ & $\oslash$& $\surd$& $\surd$ \\ \hline
Binary A & $\surd$& $\surd$& $\surd$ & $\surd$& $\surd$& $\surd$ \\ \hline
Categorical A & $\surd$& $\surd$& $\surd$ & $\surd$& $\surd$& $\surd$ \\ \hline
Single M& $\surd$& $\surd$& $\surd$ & $\surd$& $\surd$& $\surd$ \\ \hline
Multiple M& $\surd$& $\surd$& $\surd$ & $\surd$& $\surd$& $\surd$ \\ \hline
Pre-exposure Confounding & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
Post-exposure Confounding & $\oslash$ & $\oslash$& $\surd$& $\oslash$& $\oslash$& $\surd$ \\ \hline
2-way Decomposition & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
4-way Decomposition & $\surd$& $\surd$& $\surd$& $\oslash$& $\times$& $\surd$ \\ \hline
Estimation: Closed-form Parameter Function & $\surd$\tnote{*}& $\oslash$& $\oslash$& $\surd$& $\surd$& $\oslash$ \\ \hline
Estimation: Direct Imputation & $\surd$& $\surd$& $\surd$& $\oslash$& $\times$& $\surd$ \\ \hline
Inference: Delta Method\tnote{**} & $\surd$& $\oslash$& $\oslash$& $\oslash$& $\surd$& $\oslash$ \\ \hline
Inference: Bootstrapping & $\surd$& $\surd$& $\surd$& $\surd$& $\surd$& $\surd$ \\ \hline
\end{tabular}
\begin{tablenotes}\footnotesize
\item[*] Not available for multiple-mediator cases; only outputs conditional causal effects.
\item[**] Only available for closed-form parameter function estimation.
\item[$\surd$] Available.
\item[$\times$] Currently not available but applicable.
\item[$\oslash$] Neither available nor applicable.
\item[ ] rb: Regression-based Approach; wb: Weighting-based Approach; msm: Marginal; iorw: Inverse Odds Ratio Weighting Approach; ne: Natural Effect Model; g-formula: G-formula Approach.
\end{tablenotes}
\end{threeparttable}
\end{center}

# Pre-treatment Confounding

## DAG

```{r echo=F,fig.width=6,fig.height=3}
g1 <- dagitty::dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
}")

dagitty::coordinates(g1) <- list(x=c(A=0,Y=4,M=2,C=0.2),y=c(A=0,Y=0,M=-0.8,C=-1))

plot(g1)
```

In the DAG, A denotes the treatment, Y denotes the outcome, M denotes a set of mediators, and C denotes a set of pre-treatment covariates.

## Estimand

**2-way decomposition in additive scale**

$CDE=E[Y_{am}-Y_{a^*m}]$

$PNDE=E[Y_{aM_a^*}-Y_{a^*M_a^*}]$

$TNDE=E[Y_{aM_a}-Y_{a^*M_a}]$

$PNIE=E[Y_{a^*M_a}-Y_{a^*M_a^*}]$

$TNIE=E[Y_{aM_a}-Y_{aM_a^*}]$

$TE=PNDE+TNIE$

$PM=\frac{TNIE}{PNDE+TE}$


**2-way decomposition in RR scale**

$rr^{CDE}=E[Y_{am}]/E[Y_{a^*m}]$

$rr^{PNDE}=E[Y_{aM_a^*}]/E[Y_{a^*M_a^*}]$

$rr^{TNDE}=E[Y_{aM_a}]/E[Y_{a^*M_a}]$

$rr^{PNIE}=E[Y_{a^*M_a}]/E[Y_{a^*M_a^*}]$

$rr^{TNIE}=E[Y_{aM_a}]/E[Y_{aM_a^*}]$

$rr^{TE}=rr^{PNDE}\times rr^{TNIE}$

$PM=\frac{rr^{PNDE}*(rr^{TNIE}-1)}{rr^{TE}-1}$

**4-way decomposition in additive scale**

$CDE=E[Y_{am}-Y_{a^*m}]$

$INT_{ref}=PNDE-CDE$

$INT_{med}=TNIE-PNIE$

$PIE=PNIE$

$prop^{CDE}=\frac{CDE}{TE}$

$prop^{INT_{ref}}=\frac{INT_{ref}}{TE}$

$prop^{INT_{med}}=\frac{INT_{med}}{TE}$

$prop^{PIE}=\frac{PIE}{TE}$

$overall^{PM}=\frac{PNIE+INT_{med}}{TE}$

$overall^{INT}=\frac{INT_{ref}+INT_{med}}{TE}$

$overall^{PE}=\frac{INT_{ref}+INT_{med}+PIE}{TE}$

**4-way decomposition in RR scale**

$err^{CDE}=(E[Y_{am}-Y_{a^*m}])/E[Y_{a^*M_a^*}]$

$err^{INT_{ref}}=rr^{PNDE}-1-err^{CDE}$

$err^{INT_{med}}=rr^{TNIE}*rr^{PNDE}-rr^{PNDE}-rr^{PNIE}+1$

$err^{PIE}=rr^{PNIE}-1$

$err^{TE}=err^{CDE}+err^{INT_{ref}}+err^{INT_{med}}+err^{PIE}=rr^{TE}-1$

$prop^{err^{CDE}}=\frac{err^{CDE}}{err^{TE}}$

$prop^{err^{INT_{ref}}}=\frac{err^{INT_{ref}}}{err^{TE}}$

$prop^{err^{INT_{med}}}=\frac{err^{INT_{med}}}{err^{TE}}$

$prop^{err^{PIE}}=\frac{err^{PIE}}{err^{TE}}$

$overall^{PM}=\frac{err^{PIE}+err^{INT_{med}}}{err^{TE}}$

$overall^{INT}=\frac{err^{INT_{ref}}+err^{INT_{med}}}{err^{TE}}$

$overall^{PE}=\frac{err^{INT_{ref}}+err^{INT_{med}}+err^{PIE}}{err^{TE}}$

## Estimation

**Closed-form Parameter Function Estimation**

Effect estimates are calculated using regression parameters. 

**Direct Conterfacturals imputation Estimation**

1. Impute $Y_{am,i}$, $Y_{a^*m,i}$, $Y_{aMa,i}$, $Y_{a^*Ma^*,i}$, $Y_{aMa^*,i}$, and $Y_{a^*Ma,i}$ for each subject i.

2. Estimate $E[Y_{am}]$, $E[Y_{a^*m}]$, $E[Y_{aMa}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa^*}]$, and $E[Y_{a^*Ma}]$ by $\frac{\sum_{i=1}^NY_{am,i}}{n}$, $\frac{\sum_{i=1}^NY_{a^*m,i}}{n}$, $\frac{\sum_{i=1}^NY_{aMa,i}}{n}$, $\frac{\sum_{i=1}^NY_{a^*Ma^*,i}}{n}$, $\frac{\sum_{i=1}^NY_{aMa^*,i}}{n}$ and $\frac{\sum_{i=1}^NY_{a^*Ma,i}}{n}$ respectively.

3. Calculate causal effects using $E[Y_{am}]$, $E[Y_{a^*m}]$, $E[Y_{aMa}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa^*}]$, and $E[Y_{a^*Ma}]$.

## Inference 

**Delta Method**

Standard errors of effects are estimated using the standard errors of regression parameters and delta method based on the closed-form parameter function.

**Bootstrapping**

Bootstrap the data, refit the regression models, and calculate a bootstrap estimate. Repeat the bootstrapping K times and calculate the standard error of these K boostrap estimates for each estimand, which is the estimated standard error of this estimand.

## Estimation Approaches Can be Used

### Regression-based Approach

Reference:

https://www.ncbi.nlm.nih.gov/pubmed/23379553

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/

#### Estimation

**Closed-form Parameter Function Estimation**

(1) Continuous Outcome and Continuous Mediator

Fit a simple linear regression model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2'c\ (for\ continuous \ exposure)$$
$$E[M|a,c]=\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c\ (for\ binary\ or\ categorical \ exposure)$$

Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c\ (for\ continuous \ exposure)$$
$$E[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\theta_2m+\sum_{k=1}^K\theta_{3k}I\{a=k\}m+\theta_4'c\ (for\ binary\ or\ categorical \ exposure)$$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$CDE=(\theta_1+\theta_3m)(a-a^*)$

$PNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c)\}(a-a^*)$

$TNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c)\}(a-a^*)$

$PNIE=(\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*)$

$TNIE=(\theta_2\beta_1+\theta_3\beta_1a)(a-a^*)$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$CDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})m$

$PNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)$

$TNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)$

$PNIE=(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\})$

$TNIE=(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\})(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\})$

&nbsp; 

(2)Continuous Outcome and Binary Mediator

Fit a logistic regression model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2'c\ (for\ continuous \ exposure)$$
$$logitE[M|a,c]=\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c\ (for\ binary\ or\ categorical \ exposure)$$

Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c\ (for\ continuous \ exposure)$$
$$E[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\theta_2m+\sum_{k=1}^K\theta_{3k}I\{a=k\}m+\theta_4'c\ (for\ binary\ or\ categorical \ exposure)$$

&nbsp;

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$CDE=(\theta_1+\theta_3m)(a-a^*)$

$PNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)}\}(a-a^*)$

$TNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}\}(a-a^*)$

$PNIE=(\theta_2+\theta_3a^*)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$

$TNIE=(\theta_2+\theta_3a)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$CDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})m$

$PNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)}$

$TNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}$

$PNIE=(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}-\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)})$

$TNIE=(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\})(\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)}-\frac{exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)})$

&nbsp; 

(3)Continuous Outcome and Categorical Mediator

Fit a multinomial logistic regression model for the mediator:

$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\beta_{1j}a+\beta_{2j}'c, j=1,2,...,l\ (for\ continuous \ exposure)$$

$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c, j=1,2,...,l\ (for\ binary\ or\ categorical \ exposure)$$

Fit a simple linear regression model for the outcome:
$$E[Y|a,m,c]=\theta_0+\theta_1a+\sum_{j=1}^l\theta_{2j}I\{m=j\}+a\sum_{j=1}^l\theta_{3j}I\{m=j\}+\theta_4'c\ (for\ continuous \ exposure)$$
$$E[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\sum_{j=1}^l\theta_{2j}I\{m=j\}+\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=j\}I\{a=k\}+\theta_4'c$$ 
$$(for\ binary\ or\ categorical \ exposure)$$

&nbsp;

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$CDE=(\theta_1+\sum_{j=1}^l\theta_{3j}I\{m=j\})(a-a^*)$

$PNDE=\{\theta_1+\frac{\sum_{j=1}^l\theta_{3j}exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}\}(a-a^*)$

$TNDE=\{\theta_1+\frac{\sum_{j=1}^l\theta_{3j}exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}\}(a-a^*)$

$PNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

$TNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$CDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=l\}I\{a=k\}-\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=l\}I\{a^*=k\})$

$PNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+\frac{\sum_{j=1}^l(\sum_{k=1}^K\theta_{3jk}I\{a=k\}-\sum_{k=1}^K\theta_{3jk}I\{a^*=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}$

$TNDE=(\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+\frac{\sum_{j=1}^l(\sum_{k=1}^K\theta_{3jk}I\{a=k\}-\sum_{k=1}^K\theta_{3jk}I\{a^*=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}$

$PNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}$

$TNIE=\frac{\sum_{j=1}^l(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)}-\frac{\sum_{j=1}^l(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\})exp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}$

&nbsp; 

(4)Binary Outcome and Continuous Mediator

Fit a simple linear regression model for the mediator:
$$E[M|a,c]=\beta_0+\beta_1a+\beta_2'c\ (for\ continuous \ exposure)$$
$$E[M|a,c]=\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c\ (for\ binary\ or\ categorical \ exposure)$$

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c\ (for\ continuous \ exposure)$$
$$logitE[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\theta_2m+\sum_{k=1}^K\theta_{3k}I\{a=k\}m+\theta_4'c\ (for\ binary\ or\ categorical \ exposure)$$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$OR^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$

$OR^{PNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$

$OR^{TNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$

$OR^{PNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*))$

$OR^{TNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a)(a-a^*))$

$comp^{CDE}=(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))exp(\theta_2m-(\theta_2+\theta_3a^*)(\beta_0+\beta_1a^*+\beta_2'c)-0.5(\theta_2+\theta_3a^*)^2\sigma^2)$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$OR^{CDE}=exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})m)$

$OR^{PNDE}=exp((\{\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\{\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c+\theta_2\sigma^2)\}+0.5\sigma^2(\sum_{k=1}^K\theta_{3k}^2I\{a=k\}-\sum_{k=1}^K\theta_{3k}^2I\{a^*=k\}))$

$OR^{TNDE}=exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c+\theta_2\sigma^2)+0.5\sigma^2(\sum_{k=1}^K\theta_{3k}^2I\{a=k\}-\sum_{k=1}^K\theta_{3k}^2I\{a^*=k\}))$

$OR^{PNIE}=exp(\theta_2(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\})+\sum_{k=1}^K\theta_{3k}I\{a^*=k\}(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\}))$

$OR^{TNIE}=exp(\theta_2(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\})+\sum_{k=1}^K\theta_{3k}I\{a=k\}(\sum_{k=1}^K\beta_{1k}I\{a=k\}-\sum_{k=1}^K\beta_{1k}I\{a^*=k\}))$

$comp^{CDE}=(exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+\sum_{k=1}^K\theta_{3k}I\{a=k\}m)-exp(\sum_{k=1}^K\theta_{3k}I\{a^*=k\}m))exp(\theta_2m-(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)-0.5(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})^2\sigma^2)$

&nbsp; 

(5)Binary Outcome and Binary Mediator

Fit a logistic regression model for the mediator:
$$logitE[M|a,c]=\beta_0+\beta_1a+\beta_2'c\ (for\ continuous \ exposure)$$

Fit a logistic regression model for the outcome:
$$logitE[M|a,c]=\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c\ (for\ binary\ or\ categorical \ exposure)$$

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\theta_2m+\theta_3am+\theta_4'c\ (for\ continuous \ exposure)$$
$$logitE[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\theta_2m+\sum_{k=1}^K\theta_{3k}I\{a=k\}m+\theta_4'c\ (for\ binary\ or\ categorical \ exposure)$$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$OR^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$

$OR^{PNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$OR^{TNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}$

$OR^{PNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$OR^{TNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}$

$comp^{CDE}=\frac{exp(\theta_2m)(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))(1+exp(\beta_0+\beta_1a^*+\beta_2'c))}{1+exp(\beta_0+\beta_1a^*+\beta_2'c+\theta_2+\theta_3a^*)}$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$OR^{CDE}=exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{k=1}^K\theta_{3k}I\{a=k\}-\sum_{k=1}^K\theta_{3k}I\{a^*=k\})m)$

$OR^{PNDE}=\frac{exp(\sum_{k=1}^K\theta_{1k}I\{a=k\})\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}}{exp(\sum_{k=1}^K\theta_{1k}I\{a^*=k\})\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}}$

$OR^{TNDE}=\frac{exp(\sum_{k=1}^K\theta_{1k}I\{a=k\})\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}}{exp(\sum_{k=1}^K\theta_{1k}I\{a^*=k\})\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}}$

$OR^{PNIE}=\frac{\{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}}{\{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}}$

$OR^{TNIE}=\frac{\{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}}{\{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a=k\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{k=1}^K\theta_{3k}I\{a=k\}+\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)\}}$

$comp^{CDE}=\frac{exp(\theta_2m)(exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+\sum_{k=1}^K\theta_{3k}I\{a=k\}m)-exp(\sum_{k=1}^K\theta_{3k}I\{a^*=k\}m))}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c+\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})}$

$\frac{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c)}{1+exp(\beta_0+\sum_{k=1}^K\beta_{1k}I\{a^*=k\}+\beta_2'c+\theta_2+\sum_{k=1}^K\theta_{3k}I\{a^*=k\})}$

&nbsp; 

(6)Binary Outcome and Categorical Mediator

Fit a multinomial logistic regression model for the mediator:

$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\beta_{1j}a+\beta_{2j}'c, j=1,2,...,l\ (for\ continuous \ exposure)$$

$$log\frac{E[M=j|a,c]}{E[M=0|a,c]}=\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c, j=1,2,...,l\ (for\ binary\ or\ categorical \ exposure)$$

Fit a logistic regression model for the outcome:
$$logitE[Y|a,m,c]=\theta_0+\theta_1a+\sum_{j=1}^l\theta_{2j}I\{m=j\}+a\sum_{j=1}^l\theta_{3j}I\{m=j\}+\theta_4'c\ (for\ continuous \ exposure)$$
$$logitE[Y|a,m,c]=\theta_0+\sum_{k=1}^K\theta_{1k}I\{a=k\}+\sum_{j=1}^l\theta_{2j}I\{m=j\}+\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=j\}I\{a=k\}+\theta_4'c$$ 
$$(for\ binary\ or\ categorical \ exposure)$$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is continuous:

$OR^{CDE}=exp((\theta_1+\sum_{j=1}^l\theta_{3j}I\{m=j\})(a-a^*))$

$OR^{PNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$OR^{TNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}$

$OR^{PNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$OR^{TNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$

$comp^{CDE}=\frac{exp(\sum_{j=1}^l\theta_{2j}I\{m=j\})(1+\sum_{j=1}^lexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c))(exp(\theta_1(a-a^*)+a\sum_{j=1}^l\theta_{3j}I\{m=j\})-exp(a^*\sum_{j=1}^l\theta_{3j}I\{m=j\}))}{1+\sum_{j=1}^lexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

&nbsp; 

Closed-form parameter function estimators for the causal effects when the exposure is binary or categorical:

$OR^{CDE}=exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+(\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=l\}I\{a=k\}-\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{m=l\}I\{a^*=k\}))$

$OR^{PNDE}=\frac{exp(\sum_{k=1}^K\theta_{1k}I\{a=k\})\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}}{exp(\sum_{k=1}^K\theta_{1k}I\{a^*=k\})\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}}$

$OR^{TNDE}=\frac{exp(\sum_{k=1}^K\theta_{1k}I\{a=k\})\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}}{exp(\sum_{k=1}^K\theta_{1k}I\{a^*=k\})\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}}$

$OR^{PNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}}$

$OR^{TNIE}=\frac{\{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a=k\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)\}}$

$comp^{CDE}=\frac{exp(\sum_{j=1}^l\theta_{2j}I\{m=j\})(1+\sum_{j=1}^lexp(\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c))}{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}$

$\frac{(exp((\sum_{k=1}^K\theta_{1k}I\{a=k\}-\sum_{k=1}^K\theta_{1k}I\{a^*=k\})+\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{a=k\}I\{m=j\})-exp(\sum_{j=1}^l\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}I\{m=j\}))}{1+\sum_{j=1}^lexp(\theta_{2j}+\sum_{k=1}^K\theta_{3jk}I\{a^*=k\}+\beta_{0j}+\sum_{k=1}^K\beta_{1jk}I\{a^*=k\}+\beta_{2j}'c)}$

&nbsp;

**Direct Conterfacturals imputation Estimation**

1. Fit a regression model for Y on A, M and C.

2. Fit a regression model for each mediator in M on A and C.

Estimation Algorithm for $E(Y_{am})$: 

3. For each subject i, simulate $Y_{am,i}$ by the predicted value of the outcome regression model under $A=a,M=m,C=C_i$. 

4. Estimate $E[Y_{am}]$ by $\frac{\sum_{i=1}^NY_{am,i}}{n}$.

Estimation Algorithm for $E[Y_{a1Ma2}]$:

3. For each subject i, simulate $M_{a2,i}$ by the predicted values of mediator regression models under $A=a2,C=C_i$.

4. For each subject i, simulate $Y_{a1Ma2,i}$ by the predicted value of the outcome regression model under $A=a1,M=M_{a2,i},C=C_i$.

5. Estimate $E[Y_{a1Ma2}]$ by $\frac{\sum_{i=1}^NY_{a1Ma2,i}}{n}$.

#### Inference

When the estimands are estimated through closed-form parameter function estimation, their standard errors can be estimated by the delta method or bootstrapping; when the estimands are estimated through direct counterfactuals imputation estimation, their standard errors can be estimated by bootstrapping.

### Weighting-based Approach

Reference:

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/

#### Estimation

**Direct Counterfactuals Imputation Estimation**

1. Fit a regression model for Y on A, M, C.

2. For $E[Y_{a^*m}]$, estimate it by taking a weighted average of the predicted Y values for subjects with $A=a^*$ if the subjects had had mediator $M=m$ rather than their own values of mediator and each subject i is given a weight $\frac{P(A=a^*)}{P(A=a^|c_i)}$.

3. For $E[Y_{am}]$, estimate it by taking a weighted average of the predicted Y values for subjects with $A=a$ if the subjects had had mediator $M=m$ rather than their own values of mediator and each subject i is given a weight $\frac{P(A=a)}{P(A=a|c_i)}$.

4. For $E[Y_{a^*Ma^*}]$, estimate it by taking a weighted average of the subjects with $A=a^*$ and each subject i is given a weight $\frac{P(A=a^*)}{P(A=a^*|c_i)}$.

5. For $E[Y_{aMa}]$, estimate it by taking a weighted average of the subjects with $A=a$ and each subject i is given a weight $\frac{P(A=a)}{P(A=a|c_i)}$.

6. For $E[Y_{aMa^*}]$, estimate it by taking a weighted average of the predicted Y values  for subjects with $A=a^*$ if the subjects had had exposure $A=a$ rather than $A=a^*$ and each subject i is given a weight $\frac{P(A=a^*)}{P(A=a^*|c_i)}$.

7. For $E[Y_{a^*Ma}]$, estimate it by taking a weighted average of the predicted Y values for subjects with $A=a$ if the subjects had had exposure $A=a^*$ rather than $A=a$ and each subject i is given a weight $\frac{P(A=a)}{P(A=a|c_i)}$.

#### Inference

Estimate standard errors of estimands by bootstrapping.

### Inverse Odds Ratio Weighting Approach

Reference:

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3954805/

https://www.ncbi.nlm.nih.gov/pubmed/25693776

#### Estimation

**Closed-form Parameter Function Estimation**

1. Fit a regression model for A given M and C

2. Calculate weights for each subject, $w_i=\frac{f_{A|M,C}(A=0|M_i,C_i)}{f_{A|M,C}(A=A_i|M_i,C_i)}$.

3. Estimate the direct effect by a weighted regression model of Y on A and C using the weights calculated in 2. The estimated direct effect is the coefficient of A in this regression model.

4. Estimate the total effect by a regression model of Y on A and C. The estimated total effect is the coefficient of A in this regression model.

5. Calculate the indirect effect by subtracting the direct effect from the total effect.

#### Inference

Estimate standard errors of estimands by bootstrapping.

### Natural Effect Model

Incorporate the Medflex package: https://www.jstatsoft.org/article/view/v076i11

#### Estimation

1. Fit a working model for Y on A, M and C.

2. For each subject i, expand the dataset by setting $x^*$ to be the observed exposure value, setting x to enumerate all potential exposure levels, and then imputing $Y_{x,M_{x^*},i}$ using the predicted value of Y under $A=x,M=M_i,C=C_i$. If the exposure is continuous, expand the dataset by setting x to be a number of draws(defaults to 5) from the conditional distribution of A given $C_i$.

3. Fit a natural effect model for Y on x,$x^*$ and C using the expanded dataset. The natural effect model should at least reflect the structure of the working model.

4. The coefficient of x captures the natural direct effect and the coefficient of $x^*$ captures the natural indirect effect.

#### Inference

The standard errors of estimands can be estimated by the delta method or bootstrapping.

### Other Approaches

Approaches talked about later can also be used.

# Post-treatment Confounding

## DAG

```{r echo=F,fig.width=6,fig.height=4}
g1 <- dagitty::dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
    A -> L
    L -> Y
    L -> M
    C -> L
}")

dagitty::coordinates(g1) <- list(x=c(A=0,Y=4,M=2,C=0.2,L=2),y=c(A=0,Y=0,M=-0.8,C=-1,L=0.5))

plot(g1)
```

In the DAG, A denotes the treatment, Y denotes the outcome, M denotes a set of mediators and $M=(M_1, M_2,...,M_k)$, C denotes a set of pre-exposure covariates and L denotes a set of post-exposure covariates and $L=(L_1, L_2,...,L_s)$.

## Estimand

**2-way decomposition in additive scale**

$CDE=E[Y_{am}-Y_{a^*m}]$

$rPNDE=E[Y_{aG_a^*}-Y_{a^*G_a^*}]$

$rTNDE=E[Y_{aG_a}-Y_{a^*G_a}]$

$rPNIE=E[Y_{a^*G_a}-Y_{a^*G_a^*}]$

$rTNIE=E[Y_{aG_a}-Y_{aG_a^*}]$

$rTE=rPNDE+rTNIE$

$PM=\frac{rTNIE}{rPNDE+rTE}$

**2-way decomposition in RR scale**

$rr^{CDE}=E[Y_{am}]/E[Y_{a^*m}]$

$rr^{rPNDE}=E[Y_{aG_a^*}]/E[Y_{a^*G_a^*}]$

$rr^{rTNDE}=E[Y_{aG_a}]/E[Y_{a^*G_a}]$

$rr^{rPNIE}=E[Y_{a^*G_a}]/E[Y_{a^*G_a^*}]$

$rr^{rTNIE}=E[Y_{aG_a}]/E[Y_{aG_a^*}]$

$rr^{rTE}=rr^{rPNDE}\times rr^{rTNIE}$

$PM=\frac{rr^{rPNDE}*(rr^{rTNIE}-1)}{rr^{rTE}-1}$

**4-way decomposition in additive scale**

$CDE=E[Y_{am}-Y_{a^*m}]$

$rINT_{ref}=rPNDE-CDE$

$rINT_{med}=rTNIE-rPNIE$

$rPIE=rPNIE$

$prop^{CDE}=\frac{CDE}{rTE}$

$prop^{rINT_{ref}}=\frac{rINT_{ref}}{rTE}$

$prop^{rINT_{med}}=\frac{rINT_{med}}{rTE}$

$prop^{rPIE}=\frac{rPIE}{rTE}$

$overall^{PM}=\frac{rPNIE+rINT_{med}}{rTE}$

$overall^{INT}=\frac{rINT_{ref}+rINT_{med}}{rTE}$

$overall^{PE}=\frac{rINT_{ref}+rINT_{med}+rPIE}{rTE}$

**4-way decomposition in RR scale**

$err^{CDE}=(E[Y_{am}-Y_{a^*m}])/E[Y_{a^*G_a^*}]$

$err^{rINT_{ref}}=rr^{rPNDE}-1-err^{CDE}$

$err^{rINT_{med}}=rr^{rTNIE}*rr^{rPNDE}-rr^{rPNDE}-rr^{rPNIE}+1$

$err^{rPIE}=rr^{rPNIE}-1$

$err^{rTE}=err^{CDE}+err^{rINT_{ref}}+err^{rINT_{med}}+err^{rPIE}=rr^{rTE}-1$

$prop^{err^{CDE}}=\frac{err^{CDE}}{err^{rTE}}$

$prop^{err^{rINT_{ref}}}=\frac{err^{rINT_{ref}}}{err^{rTE}}$

$prop^{err^{rINT_{med}}}=\frac{err^{rINT_{med}}}{err^{rTE}}$

$prop^{err^{rPIE}}=\frac{err^{rPIE}}{err^{rTE}}$

$overall^{PM}=\frac{err^{rPIE}+err^{rINT_{med}}}{err^{rTE}}$

$overall^{INT}=\frac{err^{rINT_{ref}}+err^{rINT_{med}}}{err^{rTE}}$

$overall^{PE}=\frac{err^{rINT_{ref}}+err^{rINT_{med}}+err^{rPIE}}{err^{rTE}}$

## Estimation Approaches Can be Used

### Marginal Structual Model

Reference:

https://journals.lww.com/epidem/fulltext/2009/01000/marginal_structural_models_for_the_estimation_of.6.aspx

#### Estimation

**Direct Counterfactuals Imputation Estimation**

Estimation Algorithm for $E(Y_{am})$:

1. Fit a weighted regression model for Y on A and M where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}\frac{P(M_1=M_{1,i}|A=a_i)}{P(M_1=M_{1,i}|A=a_i,C=c_i,L=l_i)}...\frac{P(M_k=M_{k,i}|A=a_i)}{P(M_k=M_{k,i}|A=a_i,C=c_i,L=l_i)}$. Then, simulate $Y_{am}$ by the predicted value of the weighted outcome regression model under $A=a,M=m$.

2. Estimate $E(Y_{am})$ by $Y_{am}$.

Estimation Algorithm for $E(Y_{a1Ga2})$:

1. Fit a weighted regression model for each mediator $M_p,p=1,2,...,k,$ on A and C where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}$. Then, for each subject i, simulate the value of $M_{p,a2,i}$ by the predicted value of $M_{p}|A=a2,C=C_i$.

2. Fit a weighted regression model for Y on A, M and C where each subject i is given a weight $\frac{P(A=a_i)}{P(A=a_i|C=c_i)}\frac{P(M_1=M_{1,i}|A=a_i)}{P(M_1=M_{1,i}|A=a_i,C=c_i,L=l_i)}...\frac{P(M_k=M_{k,i}|A=a_i)}{P(M_k=M_{k,i}|A=a_i,C=c_i,L=l_i)}$. Then, for each subject i, simulate the potential outcome $Y_{a1Ga2,i}$ by the predicted value of $Y|A=a1,M_1=M_{1,a2,i},M_2=M_{2,a2,i},...,M_k=M_{k,a2,i},C=C_i$.

3. Estimate $E(Y_{a1Ga2})$ by $\frac{\sum_{i=1}^NY_{a1Ga2,i}}{n}$.

#### Inference

Estimate standard errors of estimands by bootstrapping.

### G-formula Approach

Reference: 

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5285457/

#### Estimation

**Direct Counterfactuals Imputation Estimation**

Estimation Algorithm for $E(Y_{am})$: 

1. Fit a model for E(Y|A,L,M,C). For subject i, simulate $Y_{am,i}$ by the predicted value of the outcome regression model under $A=a,L=L_i,M=m,C=C_i$.

2. Estimate $E(Y_{am})$ by $\frac{\sum_{i=1}^NY_{am,i}}{n}$

Estimation Algorithm for $E(Y_{a1Ga2})$:

1. Fit a regression model for each post-exposure covariate $L_q|A,C),q=1,2,...,s,$ on A and C. Then, for each subject i, simulate the value of $L_{q,a1,i}$ by the predicted value of $L_{q}|A=a1,C=C_i$.

2. Fit a regression model for each mediator $M_p,p=1,2,...,k,$ on A, C and L. Then, for each subject i, simulate the value of $M_{p,a2,i}$ by the predicted value of $M_p|A=a2,C=C_i,L=L_i$.

3. Fit a regression model for Y on A, L, M, and C. Then, for each subject i, simulate the potential outcome $Y_{a1Ga2,i}$ by the predicted value of $Y|A=a1,L_1=L_{1,a1,i},L_2=L_{2,a1,i},...,L_s=L_{s,a1,i},M_1=M_{1,a2,i},M_2=M_{2,a2,i},...,M_n=M_{k,a2,i},C=C_i$.

4. Estimate $E(Y_{a1Ga2})$ by $\frac{\sum_{i=1}^NY_{a1Ga2,i}}{n}$.

#### Inference

Estimate standard errors of estimands by bootstrapping.
