---
title: "Statistical Modeling for a Case Control Study"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Statistical Modeling for a Case Control Study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example demonstrates how to use `cmest` for a case control study. For this purpose, we simulate some data containing a continuous baseline confounder $C_1$, a binary baseline confounder $C_2$, a binary exposure $A$, a binary mediator $M$ and a binary outcome $Y$. We sample 2000 cases out of all cases and sample 2000 controls out of all controls. The true regression models for $A$, $M$ and $Y$ are:
$$logit(E(A|C_1,C_2))=0.2+0.5C_1+0.1C_2$$
$$logit(E(M|A,C_1,C_2))=1+2A+1.5C_1+0.8C_2$$
$$logit(E(Y|A,M,C_1,C_2)))=-5+0.8A-1.8M+0.5AM+0.3C_1-0.6C_2$$

```{r}
set.seed(1)
# data simulation
expit <- function(x) exp(x)/(1+exp(x))
n <- 1000000
C1 <- rnorm(n, mean = 1, sd = 0.1)
C2 <- rbinom(n, 1, 0.6)
A <- rbinom(n, 1, expit(0.2 + 0.5*C1 + 0.1*C2))
M <- rbinom(n, 1, expit(1 + 2*A + 1.5*C1 + 0.8*C2))
Y <- rbinom(n, 1, expit(-5 + 0.8*A - 1.8*M + 0.5*A*M + 0.3*C1 - 0.6*C2))
yprevalence <- sum(Y)/n
data <- data.frame(A, M, Y, C1, C2)
case_indice <- sample(which(data$Y == 1), 2000, replace = FALSE)
control_indice <- sample(which(data$Y == 0), 2000, replace = FALSE)
data <- data[c(case_indice, control_indice), ]
```

The DAG for this scientific setting is:
  
```{r}
library(CMAverse)
cmdag(outcome = "Y", exposure = "A", mediator = "M",
      basec = c("C1", "C2"), postc = NULL, node = TRUE, text_col = "white")
```

For a case control study, we set the `casecontrol` argument to be `TRUE`. It requires that either the prevalence of the case be known or the case be rare. We use the regression-based approach for illustration.

If the prevalence of the case is known, we specify it by the `yprevalence` argument. The results are:

```{r message=F,warning=F,results='hide'}
res_yprevelence <- cmest(data = data, model = "rb", casecontrol = TRUE, yprevalence = yprevalence,
                         outcome = "Y", exposure = "A",
                         mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                         mreg = list("logistic"), yreg = "logistic",
                         astar = 0, a = 1, mval = list(1), 
                         estimation = "paramfunc", inference = "delta")
```

```{r message=F,warning=F}
summary(res_yprevelence)
```

If the prevalence of the case is unknown but we know the case is rare, we set the `yrare` argument to be `TRUE`. The results are:

```{r message=F,warning=F,results='hide'}
res_yrare <- cmest(data = data, model = "rb", casecontrol = TRUE, yrare = TRUE,
                   outcome = "Y", exposure = "A",
                   mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                   mreg = list("logistic"), yreg = "logistic",
                   astar = 0, a = 1, mval = list(1),
                   estimation = "paramfunc", inference = "delta")
```

```{r message=F,warning=F}
summary(res_yrare)
```
