---
title: "Multistate approach for stochastic interventions on a time-to event mediator in the presence of competing risks"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Multistate approach for stochastic interventions on a time-to event mediator in the presence of competing risks}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4)
library(dplyr)
library(survminer)
library(mstate)
library(CMAverse)
library(foreach)
```

The DAG for this scientific setting is as follows:

```{r DAG of interest}
cmdag(outcome = "S", exposure = "A", mediator = "M",
      basec = c("C1", "C2"), postc = NULL, node = TRUE, text_col = "white")
```

An example is given to demonstrate how to use `cmest_multistate` when there is a
time-to-event mediator and a time-to-event outcome, and the estimation
of RD and SD is of interest. For this purpose, we simulate a dataset
with a binary exposure A, a time-to-event mediator M, a timeto-event
outcome S, and two baseline covariates: one continuous (C1), and one
binary (C2).

## Data simulation

We simulate data following the structure shown in the above DAG.
Specifically, we separately generate the following transitions under a
Cox proportional hazards model with the constant baseline hazard, i.e.,
assuming exponential time to event:

-   Transition 1 ($A \rightarrow M$):
    $\lambda_{01}(t|A,C1,C2) = \lambda_{01}^0(t)e^{a_1A+a_2C1+a_3C2} = 1.5 \cdot e^{=1.9A+0.2C1+0.5C2}$
-   Transition 2 ($A \rightarrow S$):
    $\lambda_{02}(t|A,C1,C2) = \lambda_{02}^0(t)e^{b_1A+b_3C1+b_4C2} = 1 \cdot e^{A-0.5C1-0.3C2}$
-   Transition 3 ($M \rightarrow S$):
    $\lambda_{12}(t|A,C1,C2) = \lambda_{12}^0(t)e^{c_1A+c_2M+c_3C1+c_4C2} = 0.8 \cdot e^{0.55A-0.15M -0.1C1-0.2C2}$

The code for generating the aforementioned data is as follows:

```{r data generation}
# set up coefficients 
# transition 1 (A to M)
a1 = -1.9 
a2 = 0.2 
a3 = 0.5 

# transition 2 (A to S)
b1 = 1 
b3 = -0.5
b4 = 0.3

# transition 3 (M to S)
c1 = 0.55
c2 = -0.15 
c3 = -0.1 
c4 = -0.2 

set.seed(8) 

# build a function to generate time-to-event data
gen_srv <- function(n, lambda, beta, X){
  X = as.matrix(X)
  beta = as.matrix(beta, ncol=1)
  time = -log(runif(n)) / (lambda * exp(X %*% beta)) 
  return(time)
}

n <- 2000 
A = sample(c(0,1),replace=TRUE, size=n, c(0.5,0.5)) #binary exposure
C1 = sample(c(0,1),replace=TRUE, size=n,c(0.6, 0.4)) #binary baseline covariate
C2 = rnorm(n, mean = 1, sd = 1) #continuous baseline covariate
id=c(1:n)
full = data.frame(id,A,C1,C2)
M = gen_srv(n=n, lambda = 1.5, beta = c(a1,a2,a3), X=data.frame(A,C1,C2)) #time to event mediator
S = gen_srv(n=n, lambda = 1, beta = c(b1,b3,b4), X=data.frame(A,C1,C2)) #time to event outcome
data = data.frame(id = c(1:n), M = M, S = S)
# indicator for event
data$ind_M = ifelse(data$M <= data$S, 1, 0)
data$ind_S = 1
data <- merge(data,full , by = "id")
# modify S distribution
trans_matrix = transMat(x = list(c(2, 3), c(3), c()), names = c("A", "M", "S"))
covs = c("A","M", "C1","C2")
pre_data = msprep(time = c(NA, "M", "S"), status = c(NA, "ind_M", "ind_S"),
                         data = data, trans = trans_matrix, keep = covs)
pre_data = expand.covs(pre_data, covs, append = TRUE, longnames = FALSE)
# resample for T < S
data_23 = pre_data[which(pre_data$trans == 3),]
data_23_tem = data.frame(id = rep(NA,dim(data_23)[1]),
                         new_y = rep(NA,dim(data_23)[1]))
paste("# to resample is ", nrow(data_23))

for(i in 1:dim(data_23)[1]){
  data_23_tem$id[i] = data_23$id[i]
  repeat {
    time_test = gen_srv(n = 1, 
                        lambda = 0.8,
                        beta = c(as.numeric(c1),
                                 c2,
                                 as.numeric(c3),
                                 as.numeric(c4)), 
                        X = data_23[i, c("A.3", "M.3", "C1.3","C2.3")])
    # exit if the condition is met
    if (time_test > data_23[i,"M.3"]) break
  }
  data_23_tem$new_y[i] = time_test
}
data_temp = merge(data, data_23_tem, by = "id", all = T)
# modify M and S
data_temp$S[which(data_temp$ind_M == 1)] = data_temp$new_y[which(data_temp$ind_M == 1)]
data_temp$M[which(data_temp$ind_M == 0)] = data_temp$S[which(data_temp$ind_M == 0)]
data_final = data_temp
data_final$A = as.factor(data_final$A) 

sc_data = data_final %>% dplyr::select(id,A,M,S,ind_M,ind_S,C1,C2)

# generate time to censoring and update event indicator
time_to_censor = runif(n, 0, 2*max(sc_data$S))
sc_data$ind_S = ifelse(sc_data$S > time_to_censor, 0, 1)
sc_data$A = factor(sc_data$A)
sc_data$C1 = factor(sc_data$C1)

```

We fit the multistate Cox proportional hazards model to the simulated
data. We can see that the regression coefficient estimates are close to
the set-up values, which is expected:

```{r check coefficient values}
## prepare the simulated data into counting process format
mstate_sc_data = msprep(time = c(NA, "M", "S"), status = c(NA, "ind_M", "ind_S"),
                         data = sc_data, trans = trans_matrix, keep = covs)
mstate_sc_data = expand.covs(mstate_sc_data, covs, append = TRUE, longnames = FALSE)

## fit mstate model to the simulated data
sc_joint_mod = coxph(Surv(Tstart, Tstop, status) ~ A.1 + A.2 + A.3 +
                         M.3 + C1.1 + C1.2 + C1.3 + C2.1 + C2.2 + C2.3 + strata(trans),
                      data = mstate_sc_data)
summary(sc_joint_mod)
```

Below, we visualize the time-to-mediator and the time-to-event
distributions by exposure group for all subjects with $C1=1$:

```{r Take a look at the simulated data}
# overlayed histogram of time to mediator and outcome distributions
hist_sc <- data.frame(value = c(sc_data$M, sc_data$S),
                 group = c(rep("M", length(sc_data$M)),
                           rep("S", length(sc_data$S))))
ggplot(hist_sc, aes(x = value, fill = group)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(title = "Overlayed Histogram of M and S",
       x = "Observed time",
       y = "Frequency") +
  scale_fill_manual(values = c("blue", "red"))

# survival curve by treatment group (A=1 vs. A=0) within strata C1=1
survival_fit_sc <- survfit(Surv(S, ind_S) ~ A, data = sc_data %>% filter(C1==1))
ggsurvplot(survival_fit_sc, data = sc_data, pval = T)
ggsurvplot(survfit(Surv(M, ind_M) ~ A, data = sc_data %>% filter(C1==1)), data = sc_data, pval = T)
```

## Demonstration

In this setting, we can use a multistate modeling approach to compute
the causal estimands of interest (RD, SD, and TE) in the presence of
semi-competing risks. We demonstrate how to achieve this using `cmest_multistate`.
In this example, we use the `cmest_multistate` function to calculate RD and SD for
time points 0.1, 0.5, 1, 2, 3, 4, 5, 6. For demonstration purposes, we
only run 2 bootstraps here. More number of bootstraps are required in
real-world settings (usually \>1000). Note this may result in longer
computation time. We have implemented parallel computing to speed-up
processing time.

```{r run multistate method on simulated data, results='hide'}
# specificy time points (s)
time_to_predict_sc <- c(0.1, 0.5, 1, 2, 3, 4, 5, 6)
# run cmest_multistate()
sc_data_result = cmest_multistate(data = sc_data, 
                                  s = time_to_predict_sc,
                                  multistate_seed = 10,
                                  exposure = 'A', mediator = 'M', outcome = 'S',
                                  yevent = "ind_S",mevent = "ind_M",
                                  basec = c("C1", "C2"),
                                  basecval = c("C1" = "1", 
                                               "C2" = as.character(mean(sc_data$C2))),
                                  astar="0", a="1", 
                                  nboot=100, EMint=F, 
                                  bh_method = "breslow")   

```

The output is a list that consists of 4 elements: 1) The model summary
of the joint multistate Cox proportional hazards model fitted on the
original dataset, 2) the point estimates of RD and SD for each of the
user-specified time points of interest on the original dataset, 3) the
summary of the bootstrapped RD, SD, and TE estimates for each of the
user-specified time point of interest, including the 2.5%, 50%, and
97.5% percentiles, and 4) the estimated RD, SD, TD for each of the
user-specified time point of interest for each bootstrap dataset. These
4 elements of the example output are extracted in order below:

```{r display cmest outputs}
sc_data_result$model_summary
sc_data_result$pt_est
sc_data_result$bootstrap_summary
sc_data_result$raw_output
```

We here provide an interpretation of RD and SD estimated from this simulated dataset:

-   TE=x can be interpreted as: The probability of surviving beyond time
    point s among the exposed group is x higher than that among the
    unexposed group, controlling for baseline covariates,

-   RD=x can be interpreted as: The probability of surviving beyond time
    point s among the exposed group is x higher than that among the
    unexposed group, controlling for baseline covariates, had the
    mediator distribution g been the fixed to that of the unexposed
    group.

-   SD=x can be interpreted as: The probability of surviving beyond time
    point s among the exposed group is estimated to increase by x, had
    the time to mediator distribution among the exposed group been
    changed to that of the unexposed group, controlling for baseline
    covariates,

## Reference

[1] Valeri, L., Proust-Lima, C., Fan, W., Chen, J. T., & Jacqmin-Gadda,
H. (2023). A multistate approach for the study of interventions on an
intermediate time-to-event in health disparities research. Statistical
Methods in Medical Research, 09622802231163331.
