---
title: "Overview of Modeling Approaches for Causal Mediation Analysis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Modeling Approaches for Causal Mediation Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document describes how the six causal mediation analysis approaches including *the regression-based approach* by [Valeri et al. (2013)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3659198/) and [VanderWeele et al. (2014)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/), 
*the weighting-based approach* by [VanderWeele et al. (2014)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4287269/), *the inverse odd-ratio weighting approach* by [Tchetgen Tchetgen (2013)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3954805/), *the natural effect model* by
[Vansteelandt et al. (2012)](https://www.degruyter.com/view/journals/em/1/1/article-p131.xml?language=en), *the marginal structural model* by [VanderWeele 
et al. (2017)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5560424/), and *the $g$-formula approach* by [Robins (1986)](https://www.sciencedirect.com/science/article/pii/0270025586900886) are implemented by the `CMAverse` package. See publications of these approaches for methodological details .

`CMAverse` currently supports a single exposure, multiple sequential mediators and a single outcome.  When multiple mediators are of interest, `CMAverse` estimates the joint mediated effect through the set of mediators. `CMAverse` also supports time varying confounders preceding the mediators. 

We categorize the causal mediation analysis approaches based on whether the approach can deal with mediator-outcome confounders affected by the exposure. Among the six approaches, only *The marginal structural model* and *the $g$-formula approach* are able to deal with mediator-outcome confounders affected by the exposure.

In this document, the outcome and the exposure are denoted as $Y$ and $A$ respectively. The set of exposure-mediator confounders, exposure-outcome confounders and mediator-outcome confounders not affected by the exposure is denoted as $C$. The set of mediators is denoted as $M$ and $M=(M_1,...,M_k)$ follows the temporal order. The set of mediator-outcome confounders affected by the exposure is denoted as $L$ and $L=(L_1,...,L_s)$ follows the temporal order. 

Since weights calculated from noncategorical variables are unstable, which hurts the performance of effect estimation and inference, weighted approaches can be implemented only for categorical exposure and mediator(s). 

# No Confounders Affected by the Exposure

## DAG

```{r echo=F,warning=F,message=F,fig.width=6,fig.height=3}
library(dagitty)
g1 <- dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
}")
coordinates(g1) <- list(x=c(A=0,Y=4,M=2,C=0.2),y=c(A=0,Y=0,M=-0.8,C=-1))
plot(g1)
```

## Estimands

For a continuous outcome, causal effects are estimated on the difference scale (summarized in table 1). For a categorical, count, or survival outcome, causal effects are estimated on the ratio scale (summarized in table 2). See [Valeri et al. (2013)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3659198/) and [VanderWeele (2015)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4220271/) for details about these effects.

```{r echo=F,warning=F,message=F}
library(knitr)
library(dplyr)
library(kableExtra)
tb1 <- data.frame(c("Controlled Direct Effect", "Pure Natural Direct Effect",
                    "Total Natural Direct Effect", "Pure Natural Indirect Effect",
                    "Total Natural Indirect Effect", "Total Effect", 
                    "Reference Interaction", "Mediated Interaction", "Proportion $CDE$",
                    "Proportion $INT_{ref}$", "Proportion $INT_{med}$", "Proportion $PNIE$", "Proportion Mediated", "Proportion Attributable to Interaction", 
                    "Proportion Eliminated"),
           c("$CDE$", "$PNDE$", "$TNDE$", "$PNIE$", "$TNIE$", "$TE$",  "$INT_{ref}$", 
             "$INT_{med}$", "$prop^{CDE}$", "$prop^{INT_{ref}}$", "$prop^{INT_{med}}$", 
             "$prop^{PNIE}$", "$PM$", "$INT$", "$PE$"),
           c("$E[Y_{am}-Y_{a^*m}]$", "$E[Y_{aM_a^*}-Y_{a^*M_a^*}]$", "$E[Y_{aM_a}-Y_{a^*M_a}]$", 
             "$E[Y_{a^*M_a}-Y_{a^*M_a^*}]$", "$E[Y_{aM_a}-Y_{aM_a^*}]$", "$PNDE+TNIE$ or $TNDE+PNIE$", 
             "$PNDE-CDE$", "$TNIE-PNIE$", "$CDE/TE$", 
             "$INT_{ref}/TE$", "$INT_{med}/TE$", "$PNIE/TE$", "$TNIE/TE$", 
             "$(INT_{ref}+INT_{med})/TE$", 
             "$(INT_{ref}+INT_{med}+PNIE)/TE$"))
colnames(tb1) <- c("Full Name", "Abbreviation", "Formula")
knitr::kable(tb1, escape = FALSE, caption = "Table 1: Causal Effects on the Difference Scale") %>%
  footnote(general = "$a$ and $a^*$ are the active and control values for $A$ respectively. $m$ is the value at which $M$ is controlled. $M_a$ denotes the counterfactual value of $M$ that would have been observed had $A$ been set to be $a$. $Y_{am}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be $m$. $Y_{aMa*}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be the counterfactual value $M_{a*}$.") %>%
  column_spec(1, width = "20em") %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = c("bordered", "striped", "hover", "responsive", "condensed"))
```

```{r echo=F,warning=F,message=F}
tb2 <- data.frame(c("Controlled Direct Effect", "Pure Natural Direct Effect",
                    "Total Natural Direct Effect", "Pure Natural Indirect Effect",
                    "Total Natural Indirect Effect", "Total Effect", 
                    "Excess Ratio due to Controlled Direct Effect", 
                    "Excess Ratio due to Reference Interaction", 
                    "Excess Ratio due to Mediated Interaction", 
                    "Excess Ratio due to Pure Natural Indirect Effect", 
                    "Proportion $ER^{CDE}$",
                    "Proportion $ER^{INT_{ref}}$", "Proportion $ER^{INT_{med}}$", 
                    "Proportion $ER^{PNIE}$", "Proportion Mediated",
                    "Proportion Attributable to Interaction", 
                    "Proportion Eliminated"),
           c("$R^{CDE}$", "$R^{PNDE}$", "$R^{TNDE}$", "$R^{PNIE}$", "$R^{TNIE}$", "$R^{TE}$", 
             "$ER^{CDE}$", "$ER^{INT_{ref}}$", "$ER^{INT_{med}}$", "$ER^{PNIE}$", 
             "$prop^{ER^{CDE}}$", 
             "$prop^{ER^{INT_{ref}}}$", "$prop^{ER^{INT_{med}}}$", 
             "$prop^{ER^{PNIE}}$", "$PM$", "$INT$", "$PE$"),
           c("$E[Y_{am}]/E[Y_{a^*m}]$", "$E[Y_{aM_a^*}]/E[Y_{a^*M_a^*}]$", 
             "$E[Y_{aM_a}]/E[Y_{a^*M_a}]$", "$E[Y_{a^*M_a}]/E[Y_{a^*M_a^*}]$", 
             "$E[Y_{aM_a}]/E[Y_{aM_a^*}]$", "$R^{PNDE}\\times R^{TNIE}$ or $R^{TNDE}\\times R^{PNIE}$", 
              "$(E[Y_{am}-Y_{a^*m}])/E[Y_{a^*M_a^*}]$",
             "$R^{PNDE}-1-ER^{CDE}$", "$R^{TNIE}*R^{PNDE}-R^{PNDE}-R^{PNIE}+1$", 
             "$R^{PNIE}-1$", "$ER^{CDE}/(R^{TE}-1)$", 
             "$ER^{INT_{ref}}/(R^{TE}-1)$", 
             "$ER^{INT_{med}}/(R^{TE}-1)$", "$ER^{PNIE}/(R^{TE}-1)$", 
             "$(R^{PNDE}*(R^{TNIE}-1))/(R^{TE}-1)$",
             "$(ER^{INT_{ref}}+ER^{INT_{med}})/(R^{TE}-1)$", 
             "$(ER^{INT_{ref}}+ER^{INT_{med}}+ER^{PNIE})/(R^{TE}-1)$"))
colnames(tb2) <- c("Full Name", "Abbreviation", "Formula")
knitr::kable(tb2, escape = FALSE, caption = "Table 2: Causal Effects on the Ratio Scale") %>%
  footnote(general = "$a$ and $a^*$ are the active and control values for $A$ respectively. $m$ is the value at which $M$ is controlled. $M_a$ denotes the counterfactual value of $M$ that would have been observed had $A$ been set to be $a$. $Y_{am}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be $m$. $Y_{aMa*}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be the counterfactual value $M_{a*}$. If $Y$ is categorical, a reference value of $Y$, denoted as $y$, is needed and $E[Y]$ represents the probability of $Y=y$.") %>%
  column_spec(1, width = "20em") %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = c("bordered", "striped", "hover", "responsive"))
```


## The Regression-based Approach

With the regression-based approach, all causal effects are estimated through either closed-form parameter function estimation or direct counterfactual imputation estimation. Standard errors of causal effects are estimated through either the delta method or bootstrapping.

### Closed-form Parameter Function Estimation

Closed-form parameter function estimation is available when there is only a single mediator, i.e., $M=M_1$. Also, `yreg` must be chosen from `linear`, `logistic`, `loglinear`, `poisson`, `quasipoisson`, `negbin`, `coxph`, `aft_exp` and `aft_weibull`. `mreg` must be chosen from `linear`, `logistic` and `multinomial`. To use `yreg = "logistic"` and `yreg = "coxph"` in closed-form parameter function estimation, the outcome must be rare. Additionally, the causal effects estimated through closed-form parameter function estimation are conditional on the value of $C$ specified by the `precval` argument. Closed-form parameter functions are summarized below.

#### Linear `yreg`, Linear `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg="linear"` and `mreg=list("linear")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a linear regression model for the mediator:
$$E[M|A,C]=\beta_0+\beta_1A+\beta_2'C$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4'C$$

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=(\theta_1+\theta_3m)(a-a^*)$
    + $PNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c)\}(a-a^*)$
    + $TNDE=\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c)\}(a-a^*)$
    + $PNIE=(\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*)$
    + $TNIE=(\theta_2\beta_1+\theta_3\beta_1a)(a-a^*)$
  
4. Calculate other effects using formulas in table 1.

#### Linear `yreg`, Linear `mreg` and Categorical Exposure

If the exposure is categorical, `yreg="linear"` and `mreg=list("linear")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a linear regression model for the mediator:
$$E[M|A,C]=\beta_0+\sum_{h=1}^H\beta_{1h}I\{A=h\}+\beta_2'C$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\theta_2M+\sum_{h=1}^H\theta_{3h}I\{A=h\}M+\theta_4'C$$

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})m$
    + $PNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)$
    + $TNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)$
    + $PNIE=(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\})$
    + $TNIE=(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\})(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\})$

4. Calculate other effects using formulas in table 1.

#### Linear `yreg`, Logistic `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg="linear"` and `mreg=list("logistic")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a logistic regression model for the mediator:
$$logitE[M|A,C]=\beta_0+\beta_1A+\beta_2'C$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4'C$$

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=(\theta_1+\theta_3m)(a-a^*)$
    + $PNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)}\}(a-a^*)$
    + $TNDE=\{\theta_1+\theta_3\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}\}(a-a^*)$
    + $PNIE=(\theta_2+\theta_3a^*)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$
    + $TNIE=(\theta_2+\theta_3a)(\frac{exp(\beta_0+\beta_1a+\beta_2'c)}{1+exp(\beta_0+\beta_1a+\beta_2'c)}-\frac{exp(\beta_0+\beta_1a^*+\beta_2'c)}{1+exp(\beta_0+\beta_1a^*+\beta_2'c)})$

4. Calculate other effects using formulas in table 1.

#### Linear `yreg`, Logistic `mreg` and Categorical Exposure

If the exposure is categorical, `yreg="linear"` and `mreg=list("logistic")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a logistic regression model for the mediator:
$$logitE[M|A,C]=\beta_0+\sum_{h=1}^H\beta_{1h}I\{A=h\}+\beta_2'C$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\theta_2M+\sum_{h=1}^H\theta_{3h}I\{A=h\}M+\theta_4'C$$

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})m$
    + $PNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)}$
    + $TNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}$
    + $PNIE=(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}-\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)})$
    + $TNIE=(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\})(\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)}-\frac{exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)})$

4. Calculate other effects using formulas in table 1.

#### Linear `yreg`, Multinomial `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg="linear"` and `mreg=list("multinomial")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a multinomial regression model for the mediator:
$$log\frac{E[M=j|A,C]}{E[M=0|A,C]}=\beta_{0j}+\beta_{1j}A+\beta_{2j}'C, j=1,2,...,J$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\theta_1A+\sum_{j=1}^J\theta_{2j}I\{M=j\}+A\sum_{j=1}^J\theta_{3j}I\{M=j\}+\theta_4'C$$

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=(\theta_1+\sum_{j=1}^J\theta_{3j}I\{m=j\})(a-a^*)$
    + $PNDE=\{\theta_1+\frac{\sum_{j=1}^J\theta_{3j}exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}\}(a-a^*)$
    + $TNDE=\{\theta_1+\frac{\sum_{j=1}^J\theta_{3j}exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}\}(a-a^*)$
    + $PNIE=\frac{\sum_{j=1}^J(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^J(\theta_{2j}+\theta_{3j}a^*)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$
    + $TNIE=\frac{\sum_{j=1}^J(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)}-\frac{\sum_{j=1}^J(\theta_{2j}+\theta_{3j}a)exp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

4. Calculate other effects using formulas in table 1.

#### Linear `yreg`, Multinomial `mreg` and Categorical Exposure

If the exposure is categorical, `yreg="linear"` and `mreg=list("multinomial")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a multinomial regression model for the mediator:
$$log\frac{E[M=j|A,C]}{E[M=0|A,C]}=\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{A=h\}+\beta_{2j}'C, j=1,2,...,J$$

2. Fit a linear regression model for the outcome:
$$E[Y|A,M,C]=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\sum_{j=1}^J\theta_{2j}I\{M=j\}+\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{M=j\}I\{A=h\}+\theta_4'C$$ 

3. Estimate $CDE$, $PNDE$, $TNDE$, $PNIE$ and $TNIE$ by the following parameter functions:
    + $CDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{m=j\}I\{a=h\}-\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{m=j\}I\{a^*=h\})$
    + $PNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+\frac{\sum_{j=1}^J(\sum_{h=1}^H\theta_{3jh}I\{a=h\}-\sum_{h=1}^H\theta_{3jh}I\{a^*=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}$
    + $TNDE=\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+\frac{\sum_{j=1}^J(\sum_{h=1}^H\theta_{3jh}I\{a=h\}-\sum_{h=1}^H\theta_{3jh}I\{a^*=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}$
    + $PNIE=\frac{\sum_{j=1}^J(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}-\frac{\sum_{j=1}^J(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}$
    + $TNIE=\frac{\sum_{j=1}^J(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)}-\frac{\sum_{j=1}^J(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\})exp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}$

4. Calculate other effects using formulas in table 1.

#### Nonlinear `yreg`, Linear `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg!="linear"` and `mreg=list("linear")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a linear regression model for the mediator:
$$E[M|A,C]=\beta_0+\beta_1A+\beta_2'C+\epsilon_M,\epsilon_M\sim N(0,\sigma^2)$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4'C$$

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$
    + $R^{PNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a^*+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$
    + $R^{TNDE}=exp(\{\theta_1+\theta_3(\beta_0+\beta_1a+\beta_2'c+\theta_2\sigma^2)\}(a-a^*)+0.5\theta_3^2\sigma^2(a^2-a^{*2}))$
    + $R^{PNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a^*)(a-a^*))$
    + $R^{TNIE}=exp((\theta_2\beta_1+\theta_3\beta_1a)(a-a^*))$
    + $ER^{CDE}=(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))exp(\theta_2m-(\theta_2+\theta_3a^*)(\beta_0+\beta_1a^*+\beta_2'c)-0.5(\theta_2+\theta_3a^*)^2\sigma^2)$

4. Calculate other effects using formulas in table 2.

#### Nonlinear `yreg`, Linear `mreg` and Categorical Exposure

If the exposure is categorical, `yreg!="linear"` and `mreg=list("linear")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a linear regression model for the mediator:
$$E[M|A,C]=\beta_0+\sum_{h=1}^H\beta_{1h}I\{A=h\}+\beta_2'C+\epsilon_M,\epsilon_M\sim N(0,\sigma^2)$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\theta_2M+\sum_{h=1}^H\theta_{3h}I\{A=h\}M+\theta_4'C$$

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})m)$
    + $R^{PNDE}=exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c+\theta_2\sigma^2)+0.5\sigma^2(\sum_{h=1}^H\theta_{3h}^2I\{a=h\}-\sum_{h=1}^H\theta_{3h}^2I\{a^*=h\}))$
    + $R^{TNDE}=exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c+\theta_2\sigma^2)+0.5\sigma^2(\sum_{h=1}^H\theta_{3h}^2I\{a=h\}-\sum_{h=1}^H\theta_{3h}^2I\{a^*=h\}))$
    + $R^{PNIE}=exp(\theta_2(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\})+\sum_{h=1}^H\theta_{3h}I\{a^*=h\}(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\}))$
    + $R^{TNIE}=exp(\theta_2(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\})+\sum_{h=1}^H\theta_{3h}I\{a=h\}(\sum_{h=1}^H\beta_{1h}I\{a=h\}-\sum_{h=1}^H\beta_{1h}I\{a^*=h\}))$
    + $ER^{CDE}=(exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+\sum_{h=1}^H\theta_{3h}I\{a=h\}m)-exp(\sum_{h=1}^H\theta_{3h}I\{a^*=h\}m))exp(\theta_2m-(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\})(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)-0.5(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\})^2\sigma^2)$

4. Calculate other effects using formulas in table 2.

#### Nonlinear `yreg`, Logistic `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg!="linear"` and `mreg=list("logistic")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a logistic regression model for the mediator:
$$logitE[M|A,C]=\beta_0+\beta_1A+\beta_2'C$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\theta_1A+\theta_2M+\theta_3AM+\theta_4'C$$

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp((\theta_1+\theta_3m)(a-a^*))$
    + $R^{PNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$
    + $R^{TNDE}=\frac{exp(\theta_1a)\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{exp(\theta_1a^*)\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}$
    + $R^{PNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a^*+\beta_0+\beta_1a^*+\beta_2'c)\}}$
    + $R^{TNIE}=\frac{\{1+exp(\beta_0+\beta_1a^*+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a+\beta_2'c)\}}{\{1+exp(\beta_0+\beta_1a+\beta_2'c)\}\{1+exp(\theta_2+\theta_3a+\beta_0+\beta_1a^*+\beta_2'c)\}}$
    + $ER^{CDE}=\frac{exp(\theta_2m)(exp(\theta_1(a-a^*)+\theta_3am)-exp(\theta_3a^*m))(1+exp(\beta_0+\beta_1a^*+\beta_2'c))}{1+exp(\beta_0+\beta_1a^*+\beta_2'c+\theta_2+\theta_3a^*)}$

4. Calculate other effects using formulas in table 2.

#### Nonlinear `yreg`, Logistic `mreg` and Categorical Exposure

If the exposure is categorical, `yreg!="linear"` and `mreg=list("logistic")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a logistic regression model for the mediator:
$$logitE[M|A,C]=\beta_0+\sum_{h=1}^H\beta_{1h}I\{A=h\}+\beta_2'C$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\theta_2M+\sum_{h=1}^H\theta_{3h}I\{A=h\}M+\theta_4'C$$

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+(\sum_{h=1}^H\theta_{3h}I\{a=h\}-\sum_{h=1}^H\theta_{3h}I\{a^*=h\})m)$
    + $R^{PNDE}=\frac{exp(\sum_{h=1}^H\theta_{1h}I\{a=h\})\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}}{exp(\sum_{h=1}^H\theta_{1h}I\{a^*=h\})\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}}$
    + $R^{TNDE}=\frac{exp(\sum_{h=1}^H\theta_{1h}I\{a=h\})\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}}{exp(\sum_{h=1}^H\theta_{1h}I\{a^*=h\})\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}}$
    + $R^{PNIE}=\frac{\{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}}{\{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}}$
    + $R^{TNIE}=\frac{\{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}}{\{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a=h\}+\beta_2'c)\}\{1+exp(\theta_2+\sum_{h=1}^H\theta_{3h}I\{a=h\}+\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c)\}}$
    + $ER^{CDE}=\frac{exp(\theta_2m)(exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+\sum_{h=1}^H\theta_{3h}I\{a=h\}m)-exp(\sum_{h=1}^H\theta_{3h}I\{a^*=h\}m))(1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c))}{1+exp(\beta_0+\sum_{h=1}^H\beta_{1h}I\{a^*=h\}+\beta_2'c+\theta_2+\sum_{h=1}^H\theta_{3h}I\{a^*=h\})}$

4. Calculate other effects using formulas in table 2.

#### Nonlinear `yreg`, Multinomial `mreg` and Noncategorical Exposure

If the exposure is not categorical, `yreg!="linear"` and `mreg=list("multinomial")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a multinomial regression model for the mediator:
$$log\frac{E[M=j|A,C]}{E[M=0|A,C]}=\beta_{0j}+\beta_{1j}A+\beta_{2j}'C, j=1,2,...,J$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\theta_1A+\sum_{j=1}^J\theta_{2j}I\{M=j\}+A\sum_{j=1}^J\theta_{3j}I\{M=j\}+\theta_4'C$$

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp((\theta_1+\sum_{j=1}^J\theta_{3j}I\{m=j\})(a-a^*))$
    + $R^{PNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$
    + $R^{TNDE}=\frac{exp(\theta_1a)\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{exp(\theta_1a^*)\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}$
    + $R^{PNIE}=\frac{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$
    + $R^{TNIE}=\frac{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)\}}$
    + $ER^{CDE}=\frac{exp(\sum_{j=1}^J\theta_{2j}I\{m=j\})(1+\sum_{j=1}^Jexp(\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c))(exp(\theta_1(a-a^*)+a\sum_{j=1}^J\theta_{3j}I\{m=j\})-exp(a^*\sum_{j=1}^J\theta_{3j}I\{m=j\}))}{1+\sum_{j=1}^Jexp(\theta_{2j}+\theta_{3j}a^*+\beta_{0j}+\beta_{1j}a^*+\beta_{2j}'c)}$

4. Calculate other effects using formulas in table 2.

#### Nonlinear `yreg`, Multinomial `mreg` and Categorical Exposure

If the exposure is categorical, `yreg!="linear"` and `mreg=list("multinomial")`, `CMAverse` estimates the causal effects by the following steps:

1. Fit a multinomial regression model for the mediator:
$$log\frac{E[M=j|A,C]}{E[M=0|A,C]}=\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{A=h\}+\beta_{2j}'C, j=1,2,...,J$$

2. Fit the specified regression model for the outcome:
$$g(E[Y|A,M,C])=\theta_0+\sum_{h=1}^H\theta_{1h}I\{A=h\}+\sum_{j=1}^J\theta_{2j}I\{M=j\}+\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{M=j\}I\{A=h\}+\theta_4'C$$ 

3. Estimate $R^{CDE}$, $R^{PNDE}$, $R^{TNDE}$, $R^{PNIE}$, $R^{TNIE}$ and $ER^{CDE}$ by the following parameter functions:
    + $R^{CDE}=exp(\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\}+\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{m=l\}I\{a=h\}-\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{m=l\}I\{a^*=h\})$
    + $R^{PNDE}=\frac{exp(\sum_{h=1}^H\theta_{1h}I\{a=h\})\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}}{exp(\sum_{h=1}^H\theta_{1h}I\{a^*=h\})\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}}$
    + $R^{TNDE}=\frac{exp(\sum_{h=1}^H\theta_{1h}I\{a=h\})\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}}{exp(\sum_{h=1}^H\theta_{1h}I\{a^*=h\})\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}}$
    + $R^{PNIE}=\frac{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}}$
    + $R^{TNIE}=\frac{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}}{\{1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a=h\}+\beta_{2j}'c)\}\{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)\}}$
    + $ER^{CDE}=\frac{exp(\sum_{j=1}^J\theta_{2j}I\{m=j\})(1+\sum_{j=1}^Jexp(\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c))(exp((\sum_{h=1}^H\theta_{1h}I\{a=h\}-\sum_{h=1}^H\theta_{1h}I\{a^*=h\})+\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{a=h\}I\{m=j\})-exp(\sum_{j=1}^J\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}I\{m=j\}))}{1+\sum_{j=1}^Jexp(\theta_{2j}+\sum_{h=1}^H\theta_{3jh}I\{a^*=h\}+\beta_{0j}+\sum_{h=1}^H\beta_{1jh}I\{a^*=h\}+\beta_{2j}'c)}$

4. Calculate other effects using formulas in table 2.

### Direct Counterfactual Imputation Estimation

`CMAverse` conducts direct counterfactual imputation estimation by the following steps:

1. Fit a regression model for $E(Y|A,M,C)$. This regression model is specified by the `yreg` argument.

2. For $p=1,...,k$, fit a regression model for the distribution of $M_p$ given $A$ and $C$. These regression models are specified by the `mreg` argument.

3. For $p=1,...,k$ and $i=1,...,n$, simulate the counterfactuals $M_{a,p,i}$ and $M_{a^*,p,i}$.
    + Simulate $M_{a,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a,C=C_i$. Denote $M_{a,i}=(M_{a,1,i},...,M_{a,k,i})^T$.
    + Simulate $M_{a^*,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a^*,C=C_i$. Denote $M_{a^*,i}=(M_{a^*,1,i},...,M_{a^*,k,i})^T$.

4. For $i=1,...,n$, obtain $E[Y_i|A=a^*,M=m,C=C_i]$, $E[Y_i|A=a,M=m,C=C_i]$, $E[Y_i|A=a^*,M=M_{a^*,i},C=C_i]$,  $E[Y_i|A=a^*,M=M_{a,i},C=C_i]$, $E[Y_i|A=a,M=M_{a^*,i},C=C_i]$ and $E[Y_i|A=a,M=M_{a,i},C=C_i]$ from the regression model in step 1.

5. Impute the counterfactuals $E[Y_{a^*m}]$, $E[Y_{am}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa}]$, $E[Y_{aMa^*}]$ and $E[Y_{a^*Ma}]$.
    + Impute $E[Y_{a^*m}]$ by taking an average of $\{E[Y_i|A=a^*,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{am}]$ by taking an average of $\{E[Y_i|A=a,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{a^*Ma^*}]$ by taking an average of $\{E[Y_i|A=a^*,M=M_{a^*,i},C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{aMa}]$ by taking an average of $\{E[Y_i|A=a,M=M_{a,i},C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{aMa^*}]$ by taking an average of $\{E[Y_i|A=a,M=M_{a^*,i},C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{a^*Ma}]$ by taking an average of $\{E[Y_i|A=a^*,M=M_{a,i},C=C_i]\}_{i=1,...,n}$.

6. Calculate causal effects with formulas in table 1 or table 2.


## The Weighting-based Approach

With the weighting-based approach, `CMAverse` estimates causal effects through direct counterfactual imputation estimation by the following steps:

1. Fit a regression model for the distribution of $E(Y|A,M,C)$. This regression model is specified by the `yreg` argument.

2. If $C$ is not empty, fit a regression model for $P(A|C)$ and obtain $P(A=A_i|C=C_i)$ for $i=1,...,n$. This regression model is specified by the `ereg` argument.

3. For $i=1,...,n$, obtain $E[Y_i|A=a^*,M=m,C=C_i]$, $E[Y_i|A=a,M=m,C=C_i]$, $E[Y_i|A=a^*,M=M_i,C=C_i]$ and $E[Y_i|A=a,M=M_i,C=C_i]$ from the regression model in step 1.

4. Impute the counterfactuals $E[Y_{a^*m}]$, $E[Y_{am}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa}]$, $E[Y_{aMa^*}]$ and $E[Y_{a^*Ma}]$.
    + Impute $E[Y_{a^*m}]$ by taking an average of $\{E[Y_i|A=a^*,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{am}]$ by taking an average of $\{E[Y_i|A=a,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{a^*Ma^*}]$ by taking a weighted average of $\{Y_i\}_{i\in \{i:A_i=a^*\}}$, and each subject i is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}$.
    + Impute $E[Y_{aMa}]$ by taking a weighted average of $\{Y_i\}_{i\in \{i:A_i=a\}}$, and each subject i is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}$.
    + Impute $E[Y_{aMa^*}]$ by taking a weighted average of $\{E[Y_i|A=a,M=M_i,C=C_i]\}_{i\in \{i:A_i=a^*\}}$, and each subject i is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}$.
    + Impute $E[Y_{a^*Ma}]$ by taking a weighted average of $\{E[Y_i|A=a^*,M=M_i,C=C_i]\}_{i\in \{i:A_i=a\}}$, and each subject i is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}$.

5. Calculate causal effects with formulas in table 1 or table 2.


## The Inverse Odds Ratio Weighting Approach

With the inverse odds ratio weighting approach, `CMAverse` estimates causal effects through direct counterfactual imputation estimation by the following steps:

1. Fit a regression model for $P(A|M,C)$ and obtain $\frac{P(A=0|M=M_i,C=C_i)}{P(A=A_i|M=M_i,C=C_i)}$ for $i=1,...,n$. This regression model is specified by the `ereg` argument. 

2. Fit a regression model for $E(Y|A,C)$. This regression model is specified by the `yreg` argument. 

3. Fit a weighted regression model for $E(Y|A,C)$ and each subject $i$ is given a weight $\frac{P(A=0|M=M_i,C=C_i)}{P(A=A_i|M=M_i,C=C_i)}$. This regression model is obtained by adding the weights to the regression model in step 2.

4. Impute the counterfactuals $E_{tot}[Y_{a}]$, $E_{tot}[Y_{a^*}]$, $E_{dir}[Y_{a}]$ and $E_{dir}[Y_{a^*}]$.
    + Impute $E_{tot}[Y_{a}]$ by taking an average of $\{E[Y_i|A=a,C=C_i]\}_{i=1,...,n}$ obtained from the regression model in step 2.
    + Impute $E_{tot}[Y_{a^*}]$ by taking an average of $\{E[Y_i|A=a^*,C=C_i]\}_{i=1,...,n}$ obtained from the regression model in step 2.
    + Impute $E_{dir}[Y_{a}]$ by taking a weighted average of $\{E[Y_i|A=a,C=C_i]\}_{i=1,...,n}$ obtained from the regression model in step 3 and each subject $i$ is given a weight $\frac{P(A=0|M=M_i,C=C_i)}{P(A=A_i|M=M_i,C=C_i)}$.
    + Impute $E_{dir}[Y_{a^*}]$ by taking a weighted average of $\{E[Y_i|A=a^*,C=C_i]\}_{i=1,...,n}$ obtained from the regression model in step 3 and each subject $i$ is given a weight $\frac{P(A=0|M=M_i,C=C_i)}{P(A=A_i|M=M_i,C=C_i)}$.

5. + For a continuous outcome, calculate $TE$ by $E_{tot}[Y_{a}]-E_{tot}[Y_{a^*}]$, calculate $PNDE$ by $E_{dir}[Y_{a}]-E_{dir}[Y_{a^*}]$ and calculate $TNIE$ by $TE-PNDE$.
   + For a categorical, count or survival outcome, calculate $TE$ by $E_{tot}[Y_{a}]/E_{tot}[Y_{a^*}]$, calculate $PNDE$ by $E_{dir}[Y_{a}]/E_{dir}[Y_{a^*}]$ and calculate $TNIE$ by $TE/PNDE$.


## The Natural Effect Model

With the natural effect model, `CMAverse` estimates causal effects through direct counterfactual imputation estimation by the following steps:

1. Fit a regression model for $E(Y|A,M,C)$. This regression model is specified by the `yreg` argument.

2. Expand the dataset using the regression model in step 1 and the `neImpute` function in the `medflex` package. The expanded dataset gives $A0$ for the direct effect and $A1$ for the indirect effect.

3. Fit the regression model in step 1 by the expanded dataset in step 2 with the exposure in the regression formula replaced by $A0$ and mediators in the regression formula replaced by $A1$, i.e.,
$Y\sim A0+A1+A0*A1+C$ if the regression formula in step 1 is $Y\sim A+M_1+M_2+A*M_1+C$.

4. For $i=1,...,n$, obtain $E[Y_i|A=a^*,M=m,C=C_i]$ and $E[Y_i|A=a,M=m,C=C_i]$ from the regression model in step 1; obtain $E[Y_i|A0=a^*,A1=a^*,C=C_i]$, $E[Y_i|A0=a^*,A1=a,C=C_i]$, $E[Y_i|A0=a,A1=a^*,C=C_i]$ and $E[Y_i|A0=a,A1=a,C=C_i]$ from the regression model in step 3.

5. Impute the counterfactuals $E[Y_{a^*m}]$, $E[Y_{am}]$, $E[Y_{a^*Ma^*}]$, $E[Y_{aMa}]$, $E[Y_{aMa^*}]$ and $E[Y_{a^*Ma}]$.
    + Impute $E[Y_{a^*m}]$ by taking an average of $\{E[Y_i|A=a^*,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{am}]$ by taking an average of $\{E[Y_i|A=a,M=m,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{a^*Ma^*}]$ by taking an average of $\{E[Y_i|A0=a^*,A1=a^*,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{aMa}]$ by taking an average of $\{E[Y_i|A0=a,A1=a,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{aMa^*}]$ by taking an average of $\{E[Y_i|A0=a,A1=a^*,C=C_i]\}_{i=1,...,n}$.
    + Impute $E[Y_{a^*Ma}]$ by taking an average of $\{E[Y_i|A0=a^*,A1=a,C=C_i]\}_{i=1,...,n}$.

6. Calculate causal effects with formulas in table 1 or table 2.


# Confounders Affected by the exposure

## DAG

```{r echo=F,warning=F,message=F,fig.width=6,fig.height=3}
g2 <- dagitty::dagitty( "dag {
    A -> Y
    A -> M -> Y
    C -> A
    C -> M
    C -> Y
    A -> L
    L -> Y
    L -> M
    C -> L
}")

dagitty::coordinates(g2) <- list(x=c(A=0,Y=4,M=2,C=0.2,L=2),y=c(A=0,Y=0,M=-0.8,C=-1,L=0.5))

plot(g2)
```

## Estimands

For a continuous outcome, causal effects are estimated on the difference scale (summarized in table 3). For a categorical, count, or survival outcome, causal effects are estimated on the ratio scale (summarized in table 4). Because of the existence of $L$, some causal effects in table 1 and table 2 are not identifiable. However, their randomized analogues are still identifiable. See [VanderWeele et al. (2014)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4214081/) for details about randomized analogues of causal effects.

```{r echo=F,warning=F,message=F}
library(knitr)
library(dplyr)
library(kableExtra)
tb3 <- data.frame(c("Controlled Direct Effect", "Randomized Analogue of $PNDE$",
                    "Randomized Analogue of $TNDE$", "Randomized Analogue of $PNIE$",
                    "Randomized Analogue of $TNIE$", "Total Effect", 
                    "Randomized Analogue of $INT_{ref}$", "Randomized Analogue of $INT_{med}$", "Proportion $CDE$",
                    "Proportion $rINT_{ref}$", "Proportion $rINT_{med}$", "Proportion $rPNIE$", "Randomized Analogue of $PM$", "Randomized Analogue of $INT$", 
                    "Randomized Analogue of $PE$"),
           c("$CDE$", "$rPNDE$", "$rTNDE$", "$rPNIE$", "$rTNIE$", "$TE$",  "$rINT_{ref}$", 
             "$rINT_{med}$", "$prop^{CDE}$", "$prop^{rINT_{ref}}$", "$prop^{rINT_{med}}$", 
             "$prop^{rPNIE}$", "$rPM$", "$rINT$", "$rPE$"),
           c("$E[Y_{am}-Y_{a^*m}]$", "$E[Y_{aG_{a^*}}-Y_{a^*G_{a^*}}]$", "$E[Y_{aG_{a}}-Y_{a^*G_{a}}]$", 
             "$E[Y_{a^*G_{a}}-Y_{a^*G_{a^*}}]$", "$E[Y_{aG_{a}}-Y_{aG_{a^*}}]$", "$rPNDE+rTNIE$ or $rTNDE+rPNIE$", 
             "$rPNDE-CDE$", "$rTNIE-rPNIE$", "$CDE/TE$", 
             "$rINT_{ref}/TE$", "$rINT_{med}/TE$", "$rPNIE/TE$", "$rTNIE/TE$", 
             "$(rINT_{ref}+rINT_{med})/TE$", 
             "$(rINT_{ref}+rINT_{med}+rPNIE)/TE$"))
colnames(tb3) <- c("Full Name", "Abbreviation", "Formula")
knitr::kable(tb3, escape = FALSE, caption = "Table 3: Causal Effects on the Difference Scale") %>%
  footnote(general = "$a$ and $a^*$ are the active and control values for $A$. $m$ is the value at which $M$ is controlled. $G_{a}$ denotes a random draw from the distribution of $M$ had $A=a$. $Y_{am}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be $m$. $Y_{aG_{a*}}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be the counterfactual value $G_{a*}$.") %>%
  column_spec(1, width = "20em") %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = c("bordered", "striped", "hover", "responsive", "condensed"))
```

```{r echo=F,warning=F,message=F}
tb4 <- data.frame(c("Controlled Direct Effect", "Randomized Analogue of $PNDE$",
                    "Randomized Analogue of $TNDE$", "Randomized Analogue of $PNIE$",
                    "Randomized Analogue of $TNIE$", "Total Effect", 
                    "Excess Ratio due to Controlled Direct Effect", 
                    "Randomized Analogue of $ER^{INT_{ref}}$", 
                    "Randomized Analogue of $ER^{INT_{med}}$", 
                    "Randomized Analogue of $ER^{PNIE}$", 
                    "Proportion $ER^{CDE}$",
                    "Proportion $rER^{INT_{ref}}$", "Proportion $rER^{INT_{med}}$", 
                    "Proportion $rER^{PNIE}$", "Randomized Analogue of $PM$", "Randomized Analogue of $INT$", 
                    "Randomized Analogue of $PE$"),
           c("$R^{CDE}$", "$rR^{PNDE}$", "$rR^{TNDE}$", "$rR^{PNIE}$", "$rR^{TNIE}$", "$R^{TE}$", 
             "$ER^{CDE}$", "$rER^{INT_{ref}}$", "$rER^{INT_{med}}$", "$rER^{PNIE}$", 
             "$prop^{ER^{CDE}}$", 
             "$prop^{rER^{INT_{ref}}}$", "$prop^{rER^{INT_{med}}}$", 
             "$prop^{rER^{PNIE}}$", "$rPM$", "$rINT$", "$rPE$"),
           c("$E[Y_{am}]/E[Y_{a^*m}]$", "$E[Y_{aG_{a^*}}]/E[Y_{a^*G_{a^*}}]$", 
             "$E[Y_{aG_{a}}]/E[Y_{a^*G_{a}}]$", "$E[Y_{a^*G_{a}}]/E[Y_{a^*G_{a^*}}]$", 
             "$E[Y_{aG_{a}}]/E[Y_{aG_{a^*}}]$", "$rR^{PNDE}\\times rR^{TNIE}$ or $rR^{TNDE}\\times rR^{PNIE}$", 
              "$(E[Y_{am}-Y_{a^*m}])/E[Y_{a^*M_a^*}]$",
             "$rR^{PNDE}-1-ER^{CDE}$", "$rR^{TNIE}*rR^{PNDE}-rR^{PNDE}-rR^{PNIE}+1$", 
             "$rR^{PNIE}-1$", "$ER^{CDE}/(rR^{TE}-1)$", 
             "$rER^{INT_{ref}}/(R^{TE}-1)$", 
             "$rER^{INT_{med}}/(R^{TE}-1)$", "$rER^{PNIE}/(R^{TE}-1)$", 
             "$(rR^{PNDE}*(rR^{TNIE}-1))/(R^{TE}-1)$",
             "$(rER^{INT_{ref}}+rER^{INT_{med}})/(R^{TE}-1)$", 
             "$(rER^{INT_{ref}}+rER^{INT_{med}}+rER^{PNIE})/(R^{TE}-1)$"))
colnames(tb4) <- c("Full Name", "Abbreviation", "Formula")
knitr::kable(tb4, escape = FALSE, caption = "Table 4: Causal Effects on the Ratio Scale") %>%
  footnote(general = "$a$ and $a^*$ are the active and control values for $A$. $m$ is the value at which $M$ is controlled. $G_{a}$ denotes a random draw from the distribution of $M$ among those with $A=a$. $Y_{am}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be $m$. $Y_{aG_{a*}}$ denotes the counterfactual value of $Y$ that would have been observed had $A$ been set to be $a$, and $M$ to be the counterfactual value $G_{a*}$. If $Y$ is categorical, a reference value of $Y$, denoted as $y$, is needed and $E[Y]$ represents the probability of $Y=y$.") %>%
  column_spec(1, width = "20em") %>%
  column_spec(2, width = "8em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = c("bordered", "striped", "hover", "responsive"))
```

## The Marginal Structural Model

With the marginal structural model, `CMAverse` estimates causal effects through direct counterfactual imputation estimation by the following steps:

1. For $p=1,...,k$, fit the regression model specified by `wmnomreg[p]` for $P(M_p|A,M_1, ...,M_{p-1})$ and obtain $P(M_p=M_{p,i}|A=A_i,M_1=M_{1,i},...,M_{p-1}=M_{p-1,i})$ for $i=1,...,n$. 

2. For $p=1,...,k$, fit the regression model specified by `wmdenomreg[p]` for $P(M_p|A,M_1, ...,M_{p-1},L,C)$ and obtain $P(M_p=M_{p,i}|A=A_i,M_1=M_{1,i},...,M_{p-1}=M_{p-1,i},L=L_i,C=C_i)$ for $i=1,...,n$. 

3. If $C$ is not empty, fit the regression model specified by `ereg` for $P(A|C)$ and obtain $P(A=A_i|C=C_i)$ for $i=1,...,n$. 

4. Add weights to the regression model specified by `yreg` for $E(Y|A,M)$ and each subject $i,i=1,...,n$ is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}\frac{P(M_1=M_{1,i}|A=A_i)}{P(M_1=M_{1,i}|A=A_i,C=C_i,L=L_i)}...\frac{P(M_k=M_{k,i}|A=A_i,M_1=M_{1,i},...,M_{k-1}=M_{k-1,i})}{P(M_k=M_{k,i}|A=A_i,M_1=M_{1,i},...,M_{k-1}=M_{k-1,i},C=C_i,L=L_i)}$. 

5. For $p=1,...,k$, add weights to the regression model specified by `mreg[p]` for the distribution of $M_p$ given $A$ and each subject $i,i=1,...,n$ is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}$.

6. For $p=1,...,k$ and $i=1,...,n$, simulate the counterfactuals $M_{a,p,i}$ and $M_{a^*,p,i}$ from the regression models in step 5.
    + Simulate $M_{a,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a$. Denote $M_{a,i}=(M_{a,1,i},...,M_{a,k,i})^T$.
    + Simulate $M_{a^*,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a^*$. Denote $M_{a^*,i}=(M_{a^*,1,i},...,M_{a^*,k,i})^T$.

7. For $i=1,...,n$, obtain $E[Y_i|A=a^*,M=m]$, $E[Y_i|A=a,M=m]$, $E[Y_i|A=a^*,M=M_{a^*,i}]$, $E[Y_i|A=a^*,M=M_{a,i}]$, $E[Y_i|A=a,M=M_{a^*,i}]$ and $E[Y_i|A=a,M=M_{a,i}]$ from the regression model in step 4.

8. Impute the counterfactuals $E[Y_{a^*m}]$, $E[Y_{am}]$, $E[Y_{a^*Ga^*}]$, $E[Y_{aGa}]$, $E[Y_{aGa^*}]$ and $E[Y_{a^*Ga}]$.
    + Impute $E[Y_{a^*m}]$ by taking a weighted average of $\{E[Y_i|A=a^*,M=m]\}_{i=1,...,n}$;
    + Impute $E[Y_{am}]$ by taking a weighted average of $\{E[Y_i|A=a,M=m]\}_{i=1,...,n}$;
    + Impute $E[Y_{a^*Ga^*}]$ by taking a weighted average of $\{E[Y_i|A=a^*,M=M_{a^*,i}]\}_{i=1,...,n}$;
    + Impute $E[Y_{aGa}]$ by taking a weighted average of $\{E[Y_i|A=a,M=M_{a,i}]\}_{i=1,...,n}$;
    + Impute $E[Y_{aGa^*}]$ by taking a weighted average of $\{E[Y_i|A=a,M=M_{a^*,i}]\}_{i=1,...,n}$;
    + Impute $E[Y_{a^*Ga}]$ by taking a weighted average of $\{E[Y_i|A=a^*,M=M_{a,i}]\}_{i=1,...,n}$,
    
   each subject $i,i=1,...,n$ is given a weight $\frac{P(A=A_i)}{P(A=A_i|C=C_i)}\frac{P(M_1=M_{1,i}|A=A_i)}{P(M_1=M_{1,i}|A=A_i,C=C_i,L=L_i)}...\frac{P(M_k=M_{k,i}|A=A_i,M_1=M_{1,i},...,M_{k-1}=M_{k-1,i})}{P(M_k=M_{k,i}|A=A_i,M_1=M_{1,i},...,M_{k-1}=M_{k-1,i},C=C_i,L=L_i)}$.

9. Calculate causal effects with formulas in table 3 or table 4.


## The G-formula Approach

With the $g$-formula approach, `CMAverse` estimates causal effects through direct counterfactual imputation estimation by the following steps:

1. For $q=1,...,s$, fit the regression model specified by `postcreg[q]` for the distribution of $L_q$ given $A$ and $C$. 

2. For $q=1,...,s$ and $i=1,...,n$, simulate the counterfactuals $L_{a,q,i}$ and $L_{a^*,q,i}$ from the regression models in step 1.
    + Simulate $L_{a,q,i}$ by randomly drawing a value from the distribution of $L_{q}$ given $A=a,C=C_i$. Denote $L_{a,i}=(L_{a,1,i},...,L_{a,s,i})^T$.
    + Simulate $L_{a^*,q,i}$ by randomly drawing a value from the distribution of $L_{q}$ given $A=a^*,C=C_i$. Denote $L_{a^*,i}=(L_{a^*,1,i},...,L_{a^*,s,i})^T$.
    
3. For $p=1,...,k$, fit the regression model specified by `mreg[p]` for the distribution of $M_p$ given $A$, $L$ and $C$. 

4. For $p=1,...,k$ and $i=1,...,n$, simulate the counterfactuals $M_{a,p,i}$ and $M_{a^*,p,i}$ from the regression models in step 3.
    + Simulate $M_{a,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a,L=L_{a,i},C=C_i$. Denote $M_{a,i}=(M_{a,1,i},...,M_{a,k,i})^T$.
    + Simulate $M_{a^*,p,i}$ by randomly drawing a value from the distribution of $M_{p}$ given $A=a^*,L=L_{a^*,i},C=C_i$. Denote $M_{a^*,i}=(M_{a^*,1,i},...,M_{a^*,k,i})^T$.

5. Obtain $\{G_{a,i}\}_{i=1,...,n}$ by randomly permuting $\{M_{a,i}\}_{i=1,...,n}$ and obtain $\{G_{a^*,i}\}_{i=1,...,n}$ by randomly permuting $\{M_{a^*,i}\}_{i=1,...,n}$.

6. Fit the regression model specified by `yreg` for $E(Y|A,M,L,C)$. 

7. For $i=1,...,n$, obtain $E[Y_i|A=a^*,M=m,L=L_{a^*,i},C=C_i]$, $E[Y_i|A=a,M=m,L=L_{a,i},C=C_i]$, $E[Y_i|A=a^*,M=G_{a^*,i},L=L_{a^*,i},C=C_i]$, $E[Y_i|A=a^*,M=G_{a,i},L=L_{a^*,i},C=C_i]$, $E[Y_i|A=a,M=G_{a^*,i},L=L_{a,i},C=C_i]$ and $E[Y_i|A=a,M=G_{a,i},L=L_{a,i},C=C_i]$ from the regression model in step 5.

8. Impute the counterfactuals $E[Y_{a^*m}]$, $E[Y_{am}]$, $E[Y_{a^*Ga^*}]$, $E[Y_{aGa}]$, $E[Y_{aGa^*}]$ and $E[Y_{a^*Ga}]$.
    + Impute $E[Y_{a^*m}]$ by taking an average of $\{E[Y_i|A=a^*,M=m,L=L_{a^*,i},C=C_i]\}_{i=1,...,n}$;
    + Impute $E[Y_{am}]$ by taking an average of $\{E[Y_i|A=a,M=m,L=L_{a,i},C=C_i]\}_{i=1,...,n}$;
    + Impute $E[Y_{a^*Ga^*}]$ by taking an average of $\{E[Y_i|A=a^*,M=G_{a^*,i},L=L_{a^*,i},C=C_i]\}_{i=1,...,n}$;
    + Impute $E[Y_{aGa}]$ by taking an average of $\{E[Y_i|A=a,M=G_{a,i},L=L_{a,i},C=C_i]\}_{i=1,...,n}$;
    + Impute $E[Y_{aGa^*}]$ by taking an average of $\{E[Y_i|A=a,M=G_{a^*,i},L=L_{a,i},C=C_i]\}_{i=1,...,n}$;
    + Impute $E[Y_{a^*Ga}]$ by taking an average of $\{E[Y_i|A=a^*,M=G_{a,i},L=L_{a^*,i},C=C_i]\}_{i=1,...,n}$,

9. Calculate causal effects with formulas in table 3 or table 4.