---
title: "CMAverse Quickstart Guide"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{CMAverse Quickstart Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We illustrate the general workflow of the `CMAverse` package by a quick example. The general workflow is:

1. Plot the DAG of causal relationships using `cmdag`.

2. Estimate causal effects and make inferences using `cmest`.

3. Conduct sensitivity analysis for unmeasured confounding and measurement error using `cmsens`.

Firstly, let's load the package.

```{r}
library(CMAverse)
```

Next, we simulate some data and plot the DAG. The simulated dataset contains a binary exposure, a binary mediator, a continuous mediator, a continuous outcome and two baseline confounders.

```{r message=F,warning=F}
set.seed(1)
n <- 100
C1 <- rnorm(n, mean = 1, sd = 1)
C2 <- rbinom(n, 1, 0.6)
pa <- exp(0.2 - 0.5*C1 + 0.1*C2)/(1 + exp(0.2 - 0.5*C1 + 0.1*C2))
A <- rbinom(n, 1, pa)
pm <- exp(1 + 0.5*A - 1.5*C1 + 0.5*C2)/ (1 + exp(1 + 0.5*A - 1.5*C1 + 0.5*C2))
M1 <- rbinom(n, 1, pm)
M2 <- rnorm(n, 2 + 0.8*A - M1 + 0.5*C1 + 2*C2, 1)
Y <- rnorm(n, mean = 0.5 + 0.4*A + 0.5*M1 + 0.6*M2 + 0.3*A*M1 + 0.2*A*M2 - 0.3*C1 + 2*C2, sd = 1)
data <- data.frame(A, M1, M2, Y, C1, C2)
```

The DAG can be plotted using `cmdag`.

```{r plot_dag,message=F,warning=F}
cmdag(outcome = "Y", exposure = "A", mediator = c("M1", "M2"), 
      basec = c("C1", "C2"), postc = NULL, node = FALSE, text_col = "black")
```

Then, we estimate causal effects using `cmest`. We use the regression-based approach for illustration. The reference values for the exposure are set to be 0 and 1. The reference values for the two mediators are set to be 1. 

```{r message=F,warning=F,results='hide'}
est <- cmest(data = data, model = "rb", outcome = "Y", exposure = "A",
                mediator = c("M1", "M2"), basec = c("C1", "C2"), EMint = TRUE,
                mreg = list("logistic", "linear"), yreg = "linear",
                astar = 0, a = 1, mval = list(1, 1),
                estimation = "imputation", inference = "bootstrap", nboot = 20)
```

Summarize and plot the results:

```{r message=F,warning=F}
summary(est)
```

```{r plot_cmest,message=F,warning=F,fig.width=8,fig.height=5}
ggcmest(est) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))
```

Lastly, we conduct sensitivity analysis for the results. Sensitivity analysis for unmeasured confounding:

```{r message=F,warning=F}
cmsens(object = est, sens = "uc")
```

Assume that $C_1$ was measured with error. Sensitivity analysis for measurement error using regression calibration with a set of assumed error standard deviations 0.1, 0.2 and 0.3:

```{r message=F,warning=F,results='hide'}
me1 <- cmsens(object = est, sens = "me", MEmethod = "rc", 
              MEvariable = "C1", MEvartype = "con", MEerror = c(0.1, 0.2, 0.3))
```

Summarize and plot the results:

```{r message=F,warning=F}
summary(me1)
```

```{r plot_cmsens_me_con,message=F,warning=F,fig.width=8,fig.height=5}
ggcmsens(me1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))
```

Then, assume that the exposure was measured with error. Sensitivity analysis for measurement error using SIMEX with two assumed misclassification matrices:

```{r message=F,warning=F,results='hide'}
me2 <- cmsens(object = est, sens = "me", MEmethod = "simex", MEvariable = "A", 
              MEvartype = "cat", B = 5,
              MEerror = list(matrix(c(0.95, 0.05, 0.05, 0.95), nrow = 2), 
                             matrix(c(0.9, 0.1, 0.1, 0.9), nrow = 2)))
```

Summarize and plot the results:

```{r message=F,warning=F}
summary(me2)
```

```{r plot_cmsens_me_cat,message=F,warning=F,fig.width=8,fig.height=6}
ggcmsens(me2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))
```
