---
title: "Statistical Modeling with a Single Mediator"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Statistical Modeling with a Single Mediator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
This example demonstrates how to use `cmest` when there is a single mediator. For this purpose, we simulate some data containing a continuous baseline confounder $C_1$, a binary baseline confounder $C_2$, a binary exposure $A$, a binary mediator $M$ and a binary outcome $Y$. The true regression models for $A$, $M$ and $Y$ are:
$$logit(E(A|C_1,C_2))=0.2+0.5C_1+0.1C_2$$
$$logit(E(M|A,C_1,C_2))=1+2A+1.5C_1+0.8C_2$$
$$logit(E(Y|A,M,C_1,C_2)))=-3-0.4A-1.2M+0.5AM+0.3C_1-0.6C_2$$
  
```{r}
library(CMAverse)
set.seed(1)
expit <- function(x) exp(x)/(1+exp(x))
n <- 10000
C1 <- rnorm(n, mean = 1, sd = 0.1)
C2 <- rbinom(n, 1, 0.6)
A <- rbinom(n, 1, expit(0.2 + 0.5*C1 + 0.1*C2))
M <- rbinom(n, 1, expit(1 + 2*A + 1.5*C1 + 0.8*C2))
Y <- rbinom(n, 1, expit(-3 - 0.4*A - 1.2*M + 0.5*A*M + 0.3*C1 - 0.6*C2))
data <- data.frame(A, M, Y, C1, C2)
```

The DAG for this scientific setting is:

```{r}
cmdag(outcome = "Y", exposure = "A", mediator = "M",
      basec = c("C1", "C2"), postc = NULL, node = TRUE, text_col = "white")
```

In this setting, we can use all of the six statistical modeling approaches. The results are shown below:

## The Regression-based Approach

### Closed-form Parameter Function Estimation and Delta Method Inference

```{r message=F,warning=F,results='hide'}
res_rb_param_delta <- cmest(data = data, model = "rb", outcome = "Y", exposure = "A",
                            mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                            mreg = list("logistic"), yreg = "logistic",
                            astar = 0, a = 1, mval = list(1), yref = 1,
                            estimation = "paramfunc", inference = "delta")
```

```{r message=F,warning=F}
summary(res_rb_param_delta)
```

### Closed-form Parameter Function Estimation and Bootstrap Inference

```{r message=F,warning=F,results='hide'}
res_rb_param_bootstrap <- cmest(data = data, model = "rb", outcome = "Y", exposure = "A",
                                mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                                mreg = list("logistic"), yreg = "logistic",
                                astar = 0, a = 1, mval = list(1), yref = 1,
                                estimation = "paramfunc", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_rb_param_bootstrap)
```

### Direct Counterfactual Imputation Estimation and Bootstrap Inference

```{r message=F,warning=F,results='hide'}
res_rb_impu_bootstrap <- cmest(data = data, model = "rb", outcome = "Y", exposure = "A",
                               mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                               mreg = list("logistic"), yreg = "logistic",
                               astar = 0, a = 1, mval = list(1), yref = 1,
                               estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_rb_impu_bootstrap)
```

## The Weighting-base Approach

```{r message=F,warning=F,results='hide'}
res_wb <- cmest(data = data, model = "wb", outcome = "Y", exposure = "A",
                mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                ereg = "logistic", yreg = "logistic",
                astar = 0, a = 1, mval = list(1), yref = 1,
                estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_wb)
```

## The Inverse Odds-ratio Weighting Approach

```{r message=F,warning=F,results='hide'}
res_iorw <- cmest(data = data, model = "iorw", outcome = "Y", exposure = "A",
                  mediator = "M", basec = c("C1", "C2"), 
                  ereg = "logistic", yreg = "logistic",
                  astar = 0, a = 1, mval = list(1), yref = 1,
                  estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_iorw)
```

## The Natural Effect Model

```{r message=F,warning=F,results='hide'}
res_ne <- cmest(data = data, model = "ne", outcome = "Y", exposure = "A",
                mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                yreg = "logistic",
                astar = 0, a = 1, mval = list(1), yref = 1,
                estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_ne)
```

## The Marginal Structural Model

```{r message=F,warning=F,results='hide'}
res_msm <- cmest(data = data, model = "msm", outcome = "Y", exposure = "A",
                 mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                 ereg = "logistic", yreg = "logistic", mreg = list("logistic"),
                 wmreg = list("logistic"),
                 astar = 0, a = 1, mval = list(1), yref = 1,
                 estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_msm)
```

## The $G$-formula Approach

```{r message=F,warning=F,results='hide'}
res_gformula <- cmest(data = data, model = "gformula", outcome = "Y", exposure = "A",
                      mediator = "M", basec = c("C1", "C2"), EMint = TRUE,
                      mreg = list("logistic"), yreg = "logistic",
                      astar = 0, a = 1, mval = list(1), yref = 1,
                      estimation = "imputation", inference = "bootstrap", nboot = 2)
```

```{r message=F,warning=F}
summary(res_gformula)
```